\begin{itemize}
\item[Trilinos]
The name of the project.  Also a Greek term which,
loosely translated means ``a string of pearls,'' 
meant to evoke an image that each Trilinos package is a pearl in its 
own right, but is even more valuable when combined with other 
packages.

\item[Package]
A self-contained collection of software in Trilinos
focused on one primary class of numerical
methods.  Also a fundamental, integral unit in the Trilinos framework.

\item[Didasko]
The tutorial of Trilinos.  Offers a quick introduction to several
Trilinos packages. It contains a printable PDF document and a
variety of well-commented browsable examples that illustrate how
to use Trilinos.

Didasko contains examples for the following Trilinos packages:
Teuchos, Epetra, EpetraExt, TriUtils, Galeri, AztecOO, IFPACK,ML,
NOX, Anasazi, TPetra.

\item[new\_package] A sample Trilinos package containing all of the
infrastructure to install a new package into the Trilinos framework.
Contains the basic directory structure, a collection of sample
configuration and build files and a sample ``Hello World'' package.
Also a website.

\item[Amesos]
The Direct Sparse Solver Package in Trilinos. The goal of
Amesos is to make AX=B as easy as it sounds, at least for direct
methods. Amesos provides clean and consistent interfaces to
several popular third party libraries.

\item[Anasazi]
An extensible and interoperable framework for large-scale eigenvalue
algorithms.The motivation for this framework is to provide a generic
interface to a collection of algorithms for solving large-scale 
eigenvalue problems.

\item[AztecOO] 
Linear solver package based on preconditioned Krylov methods.  A 
follow-on to the Aztec solver package~\cite{Aztec2.1}.  
Supports all Aztec 
interfaces and functionality, but also provides significant new 
functionality.

\item[Belos] A Greek term meaning ``arrow.'' Belos is the next
generation of iterative solvers.  Belos solvers are written using
``generic'' programming techniques.  In other words, Belos is written
using TSF abstract interfaces and therefore has no explicit dependence
on any concrete linear algebra library.  Instead, Belos solvers can be
used with any concrete linear algebra library that implements the TSF
abstract interfaces. 

\item[Claps]
Claps is a collection of domain decomposition preconditioners and
solvers.

\item[Epetra]
See description under "Petra".

\item[EpetraExt]
A set of extensions to Epetra. To allow Epetra to remain focused on
it's primary functionality as a Linear Algebra object support,
EpetraExt was created to maintain additional support for such
capabilities as transformations (permutions, sub-block views, etc.),
coloring support, partitioning (Zoltan), and I/O.
 
\item[Galeri]
Contains a suite of utilities and classes to generate a variety of
(distributed) linear systems. Galeri's functionalities are very close
to that of the MATLAB's gallery() function.

\item[Ifpack] 
Object-oriented algebraic preconditioner, compatible with 
Epetra and AztecOO.  Supports construction and use of parallel
distributed memory preconditioners such as overlapping Schwarz domain
decomposition, Jacobi scaling and local Gauss-Seidel relaxations.

\item[Isorropia]
A repartitioning/rebalancing package intended to assist with
redistributing objects such as matrices and matrix-graphs in a parallel
execution setting, to allow for more efficient computations.
Isorropia is primarily an interface to the Zoltan library, but can
be built and used with minimal capability without Zoltan. 

\item[Jpetra]
See description under "Petra".

\item[Kokkos]
A collection of the handful of sparse and dense kernels that determine
much of the performance for preconditioned Krylov methods. In particular,
it contains function class for sparse matrix vector multiplication and
triangular solves, and also for dense kernels that are not part of the
standard BLAS.  Kokkos is not intended as a user package, but to be
incorporated into other packages that need high performance kernels.

\item[Komplex] 
Complex linear equation solver using equivalent real 
formulations~\cite{DayHero2000}, built on top of Epetra and AztecOO.

\item[LOCA]
Library of continuation algorithms. A package of scalable stability 
analysis algorithms (available as part of 
the NOX nonlinear solver package). When integrated into an application code, 
LOCA enables the tracking of solution branches as a function of system 
parameters and the direct tracking of bifurcation points.

\item[Meros]
Segregated preconditioning package.  Provides scalable block
preconditioning for problems that couple simultaneous solution
variables such as Navier-Stokes problems. 

\item[ML]
Algebraic multi-level preconditioner package.  Provides scalable
preconditioning capabilities for a variety of problem classes.  

\item[Moertel]
Supplies capabilities for nonconforming mesh tying and contact
formulations in 2 and 3 dimensions using Mortar methods.

\item[MOOCHO]
MOOCHO (Multifunctional Object-Oriented arCHitecture for Optimization) is
designed to solve large-scale, equality and inequality nonlinearly
constrained, non-convex optimization problems (i.e. nonlinear programs)
using reduced-space successive quadratic programming (SQP) methods. 

\item[NOX]
A collection of nonlinear solvers, designed to be easily integrated
into an application and used with many different linear solvers.

\item[Petra]
A Greek term meaning ``foundation.''  Trilinos has three Petra 
libraries: Epetra, Tpetra and Jpetra that provide basic classes 
for constructing and manipulating matrix, graph and vector
objects.  Epetra is the current production version that is
split into two packages, one core and one extensions.

\begin{itemize}

\item[Epetra] Current C++ production implementation of the Petra
Object Model.  The ``E'' in Epetra stands for ``essential'' implying
that this version provides the most important capabilities that are
commonly needed by our target application base.  Epetra supports real,
double-precision floating point data only (no single-precision or
complex).  Epetra avoids explicit use of some of the more
advanced features of C++, including templates and the Standard
Template Library, that can be impediments to portability.

\item[Tpetra] The future C++ version of Petra, using templates and
other more advanced features of C++.  Tpetra supports arbitrary scalar
and ordinal types, and makes extensive use of advanced C++ features.

\item[Jpetra] A Java implementation of Petra, supporting real,
double-precision data.  Written in pure Java, it is designed to be
byte-code portable and can be executed across multiple compute nodes.

\end{itemize}

\item[Pliris]
An object-oriented interface to a LU solver for dense matrices on
parallel platforms.  These matrices are double precision real
matrices distributed on a parallel machine.

\item[PyTrilinos]
A set of python wrappers for selected Trilinos packages.  Allows a
python programmer to dynamically import Trilinos packages into a python
script or the python command-line interpreter, allowing the creation
and modification of Trilinos objects and the execution of Trilinos
algorithms, without the need to constantly recompile.

\item[RTOp]
RTOp (reduction/transformation operators) provides the basic mechanism
for implementing vector operations in a flexible and efficient manner.

\item[Rythmos]
A transient integrator for ordinary differential equations and
differential-algebraic equations with support for explicit, implicit,
one-step and multi-step algorithms.  Aimed at supporting operator-split
algorithms, multi-physics applications, block linear algebra, and
adjoint integration.

\item[Sacado]
A set of automatic differentiation tools for C++ applications.  Provides
templated classes for forward, reverse and Taylor mode automatic
differentiation.

\item[Stratimikos]
Contains a unified set of Thyra-based wrappers to linear solver and
preconditioner capabilities in Trilinos.  Can also be used for unified
testing of linear solvers and preconditioners.

\item[Teuchos] A collection of classes and service software that is
useful to almost all Trilinos packages.  Includes
reference-counted pointers, parameter lists, templated interfaces to
BLAS, LAPACK and traits for templates.  

\item[Thyra] A set of interfaces and supporting code that defines basic 
interoperability mechanisms between different types of numerical software. The 
foundation of all of these interfaces are the mathematical concepts of 
vectors, vector spaces, and linear operators. All other interfaces and support 
software is built on the basic operator/vector interfaces.

\item[Tpetra]
See description under "Petra".

\item[TriUtils]
A package of utilities for other Trilinos packages. 

\item[WebTrilinos]
A scientific portal; a web-based environment to use several Trilinos
packages through the web. 

\end{itemize}

