% @HEADER
% ***********************************************************************
% 
%            Trilinos: An Object-Oriented Solver Framework
%                 Copyright (2001) Sandia Corporation
% 
% Under terms of Contract DE-AC04-94AL85000, there is a non-exclusive
% license for use of this work by or on behalf of the U.S. Government.
% 
% This library is free software; you can redistribute it and/or modify
% it under the terms of the GNU Lesser General Public License as
% published by the Free Software Foundation; either version 2.1 of the
% License, or (at your option) any later version.
%  
% This library is distributed in the hope that it will be useful, but
% WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
% Lesser General Public License for more details.
%  
% You should have received a copy of the GNU Lesser General Public
% License along with this library; if not, write to the Free Software
% Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307
% USA
% Questions? Contact Michael A. Heroux (maherou@sandia.gov) 
% 
% ***********************************************************************
% @HEADER

\documentclass[10pt,letter]{article}
\usepackage{graphicx}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{amssymb}

\usepackage{times}

\setlength\oddsidemargin {-0pt}\setlength\evensidemargin{-0pt}
\setlength{\textwidth}{150mm}\setlength{\textheight}{210mm}
\headheight=12.45pt

\newcommand{\TriExe}[1]{{\tt\$\{TRILINOS\_HOME\}$\!\!\!$ /doc/ $\!\!\!$Tutorial/ $\!\!\!$#1}}
\author{Marzio Sala and Michael Heroux \\
Computational Mathematics and Algorithms Department \\
Sandia National Laboratories \\
P.O. Box 5800 \\
Albuquerque, NM 87185-1110
}

\title{Trilinos 3.1 Tutorial \\
   SAND2003-XXX \\
December 2003 \\
Unlimited Release}

\newcommand{\Trilinos}{Trilinos}
\newcommand{\TrilinosTM}{Trilinos \copyright}

\newtheorem{remark}{Remark}

\begin{document}

\maketitle

\begin{abstract}
  
  This document introduces the use of \Trilinos{}, version 3.1.
  \Trilinos{} has been written to support, in a rigorous manner, the
  solver needs of the engineering and scientific applications at Sandia
  National Laboratories.

  \medskip
  
  Aim of this manuscript is to present the basic features of some of the
  Trilinos packages. The presented material includes the definition of
  distributed matrices and vectors with Epetra, the iterative solution
  of linear system with AztecOO, incomplete factorizations with IFPACK,
  multilevel methods with ML, direct solution of linear system with
  Amesos, and iterative solution of nonlinear systems with NOX. With the
  help of several examples, some of the most important classes and
  methods are detailed to the unexperienced user.  For the most
  majority, each example is largely commented throughout the text. Other
  comments can be found in the source of each example.

  \medskip
  
  This document is a companion to the Trilinos User's
  Guide~\cite{Trilinos-Users-Guide} and Trilinos Development
  Guides~\cite{Trilinos-Dev-Guide,Trilinos-Dev-Guide-II}. Also, the
  documentation included in each of the Trilinos' packages is of
  fundamental importance.
 
\end{abstract}

\clearpage
\section*{Acknowledgments}
The authors would like to acknowledge the support of the ASCI and LDRD programs
that funded development of Trilinos.

\clearpage

\tableofcontents

\clearpage

\include{introduction}

\include{getting_started}

\include{epetra_vector}

\include{epetra_matrix}

\include{epetra_others}

\include{aztecoo}

\include{ifpack}

\include{ml}

\include{amesos}

\include{nox}

\include{triutils}

\bibliographystyle{plain}
\bibliography{../../CommonFiles/TrilinosBibliography}

\end{document}
