% @HEADER
% ***********************************************************************
% 
%            Trilinos: An Object-Oriented Solver Framework
%                 Copyright (2001) Sandia Corporation
% 
% Under terms of Contract DE-AC04-94AL85000, there is a non-exclusive
% license for use of this work by or on behalf of the U.S. Government.
% 
% This library is free software; you can redistribute it and/or modify
% it under the terms of the GNU Lesser General Public License as
% published by the Free Software Foundation; either version 2.1 of the
% License, or (at your option) any later version.
%  
% This library is distributed in the hope that it will be useful, but
% WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
% Lesser General Public License for more details.
%  
% You should have received a copy of the GNU Lesser General Public
% License along with this library; if not, write to the Free Software
% Foundation, Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307
% USA
% Questions? Contact Michael A. Heroux (maherou@sandia.gov) 
% 
% ***********************************************************************
% @HEADER

\documentclass[12pt,relax]{SANDreport}
\usepackage{graphicx}
\usepackage{amsmath,amsfonts,amsthm}
\usepackage{amssymb}

\usepackage{times}

\def\choicebox#1#2{\noindent$\hphantom{th}$\parbox[t]{1.8in}{\sf
#1}\parbox[t]{4.5in}{#2}\\[0.8em]}

\newcommand{\TriExe}[1]{{\tt\$\{TRILINOS\_HOME\}/doc/Tutorial/#1}}
\author{Marzio Sala, Michael Heroux, David Day \\
Computational Mathematics and Algorithms Department \\
Sandia National Laboratories \\
P.O. Box 5800 \\
Albuquerque, NM 87185-1110
}

\title{Trilinos 4.0 Tutorial}
\SANDnum{SAND2004-XXX}
\SANDauthor{
Marzio Sala, Michael Heroux, David Day}

\SANDprintDate{January 2004}
\SANDreleaseType{Unlimited Release}

\newcommand{\Trilinos}{Trilinos}
\newcommand{\TrilinosTM}{Trilinos \copyright}

\newtheorem{remark}{Remark}

\begin{document}

\maketitle

\begin{abstract}
  
  The Trilinos Project is an effort to facilitate the design,
  development, integration and ongoing support of mathematical software
  libraries.  The goal of the Trilinos Project is to develop parallel
  solver algorithms and libraries within an object-oriented software
  framework for the solution of large-scale, complex multiphysics
  engineering and scientific applications. The emphasis is on developing
  robust, scalable algorithms in a software framework, using abstract
  interfaces for flexible interoperability of components while providing
  a full-featured set of concrete classes that implement all the
  abstract interfaces.

  \medskip
  
  This document introduces the use of \Trilinos{}, version 4.0.
  \Trilinos{} has been written to support, in a rigorous manner, the
  solver needs of the engineering and scientific (massively parallel)
  applications at the Sandia National Laboratories.

  \medskip
  
  The aim of this manuscript is to present the basic features of
  Trilinos. The presented material includes the definition of
  distributed matrices and vectors with Epetra, the iterative solution
  of linear systems with AztecOO, incomplete factorizations with IFPACK,
  multilevel and domain decomposition preconditioners with ML, direct
  solution of linear system with Amesos, eigenvalue and eigenvector
  computations with Anasazi, iterative solution of nonlinear systems
  with NOX.
  
  The tutorial is a self-contained introduction, intented to help users
  effectively apply the appropriate Trilinos package to their
  applications. In each basic, each examples are presented that are fit
  to be imitated. 

  \medskip
  
  This document is a companion to the Trilinos User's
  Guide~\cite{Trilinos-Users-Guide} and Trilinos Development
  Guides~\cite{Trilinos-Dev-Guide,Trilinos-Dev-Guide-II}. Please note
  that the documentation included in each of the Trilinos' packages is
  of fundamental importance.
 
\end{abstract}

\clearpage
\section*{Acknowledgments}
The authors would like to acknowledge the support of the ASCI and LDRD programs
that funded development of Trilinos.

\medskip

The developers of each Trilinos' package are acknowledged for providing
excellent documentation and examples, without which this document would
have never been possible.


\clearpage

\SANDmain

\tableofcontents

\clearpage

\include{introduction}

%\clearpage
%\newpage
%\include{getting_started}

\clearpage
\newpage
\include{epetra_vector}

\clearpage
\newpage
\include{epetra_matrix}

\clearpage
\newpage
\include{epetra_others}

\clearpage
\newpage
\include{triutils}

\clearpage
\newpage
\include{aztecoo}

\clearpage
\newpage
\include{ifpack}

\clearpage
\newpage
\include{teuchos}

\clearpage
\newpage
\include{ml}

\clearpage
\newpage
\include{amesos}

\clearpage
\newpage
\include{anasazi}

\clearpage
\newpage
\include{nox}

\bibliographystyle{plain}
\bibliography{../../CommonFiles/TrilinosBibliography,tutorial_biblio}

\end{document}
