 \chapter{Introduction}

\exo{} is the successor of the widely used finite element 
(FE) data file format EXODUS [1] (henceforth referred to as EXODUS I)
developed by Mills-Curran and Flanagan. It continues the concept of a
common database for multiple application codes (mesh generators,
analysis codes, visualization software, etc.) rather than
code-specific utilities, affording flexibility and robustness for both
the application code developer and application code user. By using the
\exo{} data model, a user inherits the flexibility of using a large
array of application codes (including vendor-supplied codes) which
access this common data file directly or via translators.

The uses of the \exo{} data model include the following:

\begin{itemize}
 \item problem definition -- mesh generation, specification 
of locations of boundary conditions and load application, specification 
of material types.
\item simulation -- model input and results output.
\item visualization -- model verification, results postprocessing, 
data interrogation, and analysis tracking.
\end{itemize}

\section{License and Availability}
The \exo{} library is licensed under the BSD open source license.

\begin{quote}
Copyright~\copyright{} 2005 National Technology & Engineering Solutions
of Sandia, LLC (NTESS).  Under the terms of Contract DE-NA0003525 with
NTESS, the U.S. Government retains certain rights in this software.

Redistribution and use in source and binary forms, with or without
modification, are permitted provided that the following conditions are
met:

\begin{itemize}
\item Redistributions of source code must retain the above copyright
      notice, this list of conditions and the following disclaimer.

\item Redistributions in binary form must reproduce the above
      copyright notice, this list of conditions and the following
      disclaimer in the documentation and/or other materials provided
      with the distribution.

\item Neither the name of NTESS nor the names of its
      contributors may be used to endorse or promote products derived
      from this software without specific prior written permission.
\end{itemize}

THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
"AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
OWNER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY
THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT
(INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE
OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
\end{quote}

The \exo{} library source code is available on GitHub
\url{https://github.com/gsjaardema/seacas.git}

For bug reports, documentation errors, and enhancement suggestions,
contact:
\begin{itemize}
   \item Gregory D. Sjaardema
   \item PHONE: (505) 844-2701
   \item EMAIL: \url{mailto:gdsjaar@sandia.gov}
   \item EMAIL: \url{mailto:gsjaardema@gmail.com}
\end{itemize}
