\section{Epetra\_CrsSingletonFilter class}
 The Epetra\_CrsSingletonFilter class takes an existing Epetra\_LinearProblem object,
analyzes it structure and explicitly eliminates singleton rows and columns from the matrix and
appropriately modifies the RHS and LHS of the linear problem.  The result of this process is a
reduced system of equations that is itself an Epetra\_LinearProblem object.  The reduced system
can then be solved using any solver that is understands an Epetra\_LinearProblem.  The 
solution for the full system is obtained by calling ComputeFullSolution().
    
Singleton rows are defined to be rows that have a single nonzero entry in the matrix.
The equation associated with this row can be explicitly eliminated because it involved
 only one variable.  For example if row i has a single nonzero value in column j, call 
it A(i,j), we can explicitly solve for x(j) = b(i)/A(i,j), where b(i) is the ith entry 
of the RHS and x(j) is the jth entry of the LHS.

Singleton columns are defined to be columns that have a single nonzero entry in the matrix.  The variable associated with this column is fully dependent, meaning that the solution for all other variables does not depend on it.  If this entry is A(i,j) then the ith row and jth column can be removed from the system and x(j) can be solved after the solution for all other variables is determined.

By removing singleton rows and columns, we can often produce a reduced system that is smaller and far less dense, and in general having better numerical properties.

The basic procedure for using this class is as follows:
\begin{EpetraItemize}
\item Construct full problem: 
      Construct and Epetra\_LinearProblem containing the "full" matrix, RHS and LHS.  
      This is done outside of Epetra\_CrsSingletonFilter class.  Presumably, you have 
      some reason to believe that this system may contain singletons.
\item Construct an Epetra\_CrsSingletonFilter instance:  Constructor needs no arguments.
\item Analyze matrix: 
      Invoke the Analyze() method, passing in the Epetra\_RowMatrix object from your full 
      linear problem mentioned in the first step above.
\item Go/No Go decision to construct reduced problem: 
      Query the results of the Analyze method using the SingletonsDetected() method.  
      This method returns "true" if there were singletons found in the matrix.  You can 
      also query any of the other methods
      in the Filter Statistics section to determine if you want to proceed with the
      construction of the reduced system.
\item Construct reduced problem: 
     If, in the previous step, you determine that you want to proceed with the
     construction of the reduced problem, you should next call the 
     ConstructReducedProblem() method, passing in the full linear problem object from the first
     step.  This method will use the information from the Analyze() method to construct a
     reduce problem that has
     explicitly eliminated the singleton rows, solved for the corresponding LHS values and
     updated the RHS.  This 
     step will also remove singleton columns from the reduced system.  Once the solution
     of the reduced problem is
     is computed (via any solver that understands an Epetra\_LinearProblem), you should
     call the ComputeFullSolution()
     method to compute the LHS values assocaited with the singleton columns.
\item Solve reduced problem: 
     Obtain a pointer to the reduced problem using the
     ReducedProblem() method.  Using the solver of your choice, solve the reduced system.
\item Compute solution to full problem:  
     Once the solution the reduced problem is
     determined, the ComputeFullSolution()
     method will place the reduced solution values into the appropriate locations of the
     full solution LHS and then
     compute the values associated with column singletons.  At this point, you have a
     complete solution to the original
     full problem.
\item Solve a subsequent full problem that differs from the original problem only in
values: It is often the case that the
     structure of a problem will be the same for a sequence of linear problems.  In this
     case, the UpdateReducedProblem(

\end{EpetraItemize}
