
\section{Introduction to {}\texttt{Ref\-Count\-Ptr<>}}
\label{rcp:sec:intro}
%

Described herein is the templated class {}\texttt{Ref\-Count\-Ptr<>}
(which exists in the namespace {}\texttt{Teuchos}) and a set of global
template functions for implementing automatic garbage collection in
C++ using smart reference-counted pointers.  This document builds on
the material in the beginner's guide
{}\cite{ref:RefCountPtrBeginnersGuide}.  The beginner's guide is
considered an integral part of this document and this document is
incomplete without it.

This discussion assumes that the reader is knowledgeable of C++ and
has read books such as {}\cite{ref:stroustrup_2000} and
{}\cite{ref:meyers_1994}.  While understanding and appreciating the
advanced features of {}\texttt{Ref\-Count\-Ptr<>} requires advanced C++
knowledge and experience, the basic use (as described in the
beginner's guide {}\cite{ref:RefCountPtrBeginnersGuide}) requires only
basic C++ skills.  By ``coding by example'' even beginner C++
programmers can immediately realize the benefits of
{}\texttt{Ref\-Count\-Ptr<>}.

The commandments stated in Appendix B in the beginner's guide
{}\cite{ref:RefCountPtrBeginnersGuide} are represented here in the
context of a wider discussion.

The following class hierarchy is used to demonstrate this smart
pointer design.

{\scriptsize\begin{verbatim}
class A { public: virtual ~A(){} A& operator=(const A&){} virtual void f(){} };
class B1 : virtual public A {};
class B2 : virtual public A {};
class C : virtual public B1, virtual public B2 {};
class D {};
class E : public D {};
\end{verbatim}

}Note that the classes {}\texttt{A}, {}\texttt{B1}, {}\texttt{B2} and
{}\texttt{C} are polymorphic (with multiple inheritance using virtual
base classes) while the classes {}\texttt{D} and {}\texttt{E} are not.
In the following description, all of the code examples are written as
though there was a {}\texttt{using namespace Teuchos;} declaration is
the current scope or as if all of the code resided in the
{}\texttt{Teuchos} namespace.

The public C++ declarations for all of the code discussed here are
shown in Appendix A in the beginner's guide
{}\cite{ref:RefCountPtrBeginnersGuide}.  A short quick-start and
reference for {}\texttt{Ref\-Count\-Ptr<>} is contained in Appendix B in
the beginner's guide {}\cite{ref:RefCountPtrBeginnersGuide}.

A word of caution; note that the design of the reference-counted smart
pointer class discussed in Item 29 of {}\cite{ref:meyers_1996} in the
subsection ``Adding Reference Counting to Existing Classes'' has a
purpose very similar to {}\texttt{Ref\-Count\-Ptr<>} but the design of
{}\texttt{Ref\-Count\-Ptr<>} is very different (see Appendix
{}\ref{rcp:apdx:design}).  Since many of the issues involved in using
smart pointers and reference counting discussed Meyers
{}\cite{ref:meyers_1996} are the same as with {}\texttt{Ref\-Count\-Ptr<>}
this reference is worthy of study.  However, the exact designs used in
this reference are not used in {}\texttt{Ref\-Count\-Ptr<>} since they
lead to less flexible software.

%
\subsection{Initializing {}\texttt{RefCountPtr} objects}
\label{rcp:sec:init-rcp-objects}
%

A smart reference-counted pointer to a dynamically allocated object
(of type {}\texttt{A} for instance) is initialized as follows

{\scriptsize\begin{verbatim}
RefCountPtr<A>             a_ptr   = rcp(new A); // A       *       a_ptr   = new A;
RefCountPtr<const A>       ca_ptr  = rcp(new A); // const A *       ca_ptr  = new A;
const RefCountPtr<A>       a_cptr  = rcp(new A); // A       * const a_cptr  = new A;
const RefCountPtr<const A> ca_cptr = rcp(new A); // const A * const ca_cptr = new A;
\end{verbatim}}

The above code shows how all the various combinations for
{}\texttt{const} or non-{}\texttt{const} pointers to {}\texttt{const}
or non-{}\texttt{const} objects are expressed (this is explained in
more detail in Item 28 in {}\cite{ref:meyers_1996}).  There is no
automatic conversion from raw pointers to smart pointers for many good
reasons (again, see Item 28 in {}\cite{ref:meyers_1996}).  To allow such
an implicit conversion almost guarantees memory
allocation/deallocation errors will occur.  Therefore, the following
templated function

{\scriptsize\begin{verbatim}
template<class T> RefCountPtr<T> rcp(T* p, bool owns_mem = true );
\end{verbatim}}

{}\noindent{}must be used to initialize a smart pointer object given a
raw pointer.  The reason for this is somewhat complicated but it is
related to multiple inheritance and/or virtual base classes (see
Appendix {}\ref{rcp:apdx:mivbc}).

\begin{commandment}\label{rcp:cmd:rcp-new}
Thou shall put a pointer for an object allocated with operator
{}\texttt{new} into a {}\texttt{Ref\-Count\-Ptr<>} object only once.  The
best way to insure this is to call operator {}\texttt{new} directly in
the call to {}\texttt{rcp(...)} to create a dynamically allocated
object that is to be managed by a {}\texttt{Ref\-Count\-Ptr<>} object.
\end{commandment}

The only case where the template function {}\texttt{rcp<>(...)} should
not be used to create a smart pointer is the initialization to
{}\texttt{NULL}.  One way to initialize a null smart pointer is to
just use the default constructor as

{\scriptsize\begin{verbatim}
RefCountPtr<A> a_ptr;   // A a_ptr = NULL:
\end{verbatim}}

One of the big advantages of using {}\texttt{Ref\-Count\-Ptr<>} over
raw pointers (independent of the reference counting and garbage
collection features) is this default initialization to
{}\texttt{NULL}.  To be more explicit about the initialization to
{}\texttt{NULL} (and for passing a null pointer into a function), a
special constructor is included which takes the enumeration value
{}\texttt{null}.  For example

{\scriptsize\begin{verbatim}
RefCountPtr<A> a_ptr = null; // i.e. A a_ptr = NULL:
\end{verbatim}}

\begin{commandment}
Thou shall only create a {}\texttt{NULL} {}\texttt{Ref\-Count\-Ptr<>}
object by using the default constructor or by using the
{}\texttt{null} enum (and its associated special constructor).  Trying
to assign to {}\texttt{NULL} or {}\texttt{0} will not compile.
\end{commandment}

Another situation that commonly occurs is when an object is allocated
on the stack or statically, but must be passed to a client that only
accepts a {}\texttt{Ref\-Count\-Ptr<>} object.  For example, suppose that
there is a class {}\texttt{Foo} that is declared as follows

{\scriptsize\begin{verbatim}
class Foo {
public:
    Foo( const RefCountPtr<A>& a ) : a_(a) {}
    void f() { a_->f(); }
private:
    RefCountPtr<A>  a_;
};
\end{verbatim}}

The following code fragment shows how to allocate an object of concrete
type {}\texttt{C} on the stack and then pass this object into the constructor
for {}\texttt{Foo}

{\scriptsize\begin{verbatim}
C c;
Foo foo(rcp(&c,false));
\end{verbatim}}

In this example, the templated function {}\texttt{rcp(...)} is used
with the parameter {}\texttt{owns\_mem} set to {}\texttt{false}.  In
this example, when the {}\texttt{foo} object goes out of scope (and
its {}\texttt{Foo::a\_} member is deleted), {}\texttt{delete} will not
be called on the underlying pointer and no memory error will occur.
For more specialized types of memory deallocation control see Section
{}\ref{rcp:sec:dealloc}.

\begin{commandment}
Thou shall only pass a raw pointer for an object that is
\underline{not} allocated by {}\texttt{new} (e.g.~allocated on the stack) into a
{}\texttt{Ref\-Count\-Ptr<>} object by using the templated function
{}\texttt{rcp<T>(T* p, bool owns\_mem)} and setting {}\texttt{owns\_mem} to
{}\texttt{false}.
\end{commandment}

%
\subsection{Accessing the underlying reference-counted object}
%

Through the magic of the overloaded member operator functions
{}\texttt{RefCountPtr<T>::operator->()} and
{}\texttt{RefCountPtr<T>::operator*()}, the underlying object being reference
counted can be accessed exactly the same as for raw pointers such as
follows

{\scriptsize\begin{verbatim}
a_ptr->f();
(*a_ptr).f();
\end{verbatim}}

Therefore, using a smart reference-counted pointer is very similar to
using a raw C++ pointer.  The use of smart pointers is very similar
to raw pointers as shown below.

{\scriptsize\begin{verbatim}
RefCountPtr<A>
  a_ptr1     = rcp(new A),   // Initialize them from a raw pointer from new
  a_ptr2     = rcp(new A);   // ""
A *ra_ptr1   = new A,        // ""
  *ra_ptr2   = new A;        // ""
a_ptr1       = rcp(ra_ptr2); // Assign from a raw pointer (only do this once!)
a_ptr2       = a_ptr1;       // Assign one smart pointer to another
a_ptr1       = rcp(ra_ptr1); // Assign from a raw pointer (only do this once!)
a_ptr1->f();                 // Access a member using ->
ra_ptr1->f();                // ""
*a_ptr1      = *a_ptr2;      // Dereference the objects and assign
*ra_ptr1     = *ra_ptr2;     // "" 
\end{verbatim}}

What makes smart pointers different however is that the above piece of code
does not create any memory leaks that would have otherwise occurred
if {}\texttt{a\_ptr1} and {}\texttt{a\_ptr2} where raw C++ pointers.

However, these smart reference-counted pointers can not be used
everywhere a raw pointer can.  {}\texttt{Ref\-Count\-Ptr<>}
purposefully does not support all raw pointer syntax.  The goal of
{}\texttt{Ref\-Count\-Ptr<>} is not to hide the fact that smart
pointers are being used over raw pointers.  Therefore, developers
should not expect (or even want) to simply change the data type of raw
pointers from {}\texttt{T*} to {}\texttt{Ref\-Count\-Ptr<T>} and then
have a large body of code simply recompile without a wimper from the
compiler. For instance, the following statements will not compile

{\scriptsize\begin{verbatim}
a_ptr1++;           // Error, pointer arithmetic ++, --, +, - etc. not defined!
a_ptr1 == ra_ptr1;  // Error, comparision operators ==, !=, <=, >= etc. not defined!
\end{verbatim}}

Since a smart reference-counted pointer can not be used everywhere a
raw C++ pointer can, there is a means for getting at the raw C++
pointer with the {}\texttt{RefCountPtr<T>::get()} function (same as
with {}\texttt{std::auto\_ptr<T>::get()}).  For example, to check if
two {}\texttt{Ref\-Count\-Ptr<>} objects contain pointers to the same
underlying object one would check:

{\scriptsize\begin{verbatim}
if( a_ptr1.get() == a_ptr2.get() )
    std::cout << "a_ptr1 and a_ptr2 point to the same object\n";
\end{verbatim}}

%
\subsection{Conversion of {}\texttt{Ref\-Count\-Ptr<>} objects}
%

The existing C++ conversion rules for raw C++ pointers ((e.g.~from
derived to base classes, from non-{}\texttt{const} to
{}\texttt{const})) are supported by {}\texttt{Ref\-Count\-Ptr<>} for
the conversion of smart pointers.  For example, the compiler will
implicitly cast a raw C++ pointer up its type's inheritance hierarchy
(e.g.~{}\texttt{B1} $\rightarrow$ {}\texttt{A} or {}\texttt{E}
$\rightarrow$ {}\texttt{D}).  For smart reference-counted pointers,
these implicit conversions are just as easy thanks to a templated copy
constructor (see Figure {}\ref{rcp:fig:decl} and the discussion in
Item 28 in {}\cite{{}\cite{ref:meyers_1996}}).  For example, the below
code compiles and runs just fine

{\scriptsize\begin{verbatim}
RefCountPtr<C>  c_ptr  = rcp(new C);
RefCountPtr<A>  a_ptr  = c_ptr;
RefCountPtr<B1> b1_ptr = c_ptr;
RefCountPtr<D>  d_ptr  = rcp(new E);
\end{verbatim}}

To perform other non-implicit type conversions with pointers such as
{}\texttt{static\-\_cast<>}, {}\texttt{const\-\_cast<>} and
{}\texttt{dynamic\-\_cast<>}, the namespace {}\texttt{Teuchos}
contains the nonmember template functions
{}\texttt{rcp\-\_static\-\_cast<>(...)},
{}\texttt{rcp\-\_const\-\_cast<>(...)} and
{}\texttt{rcp\-\_dynamic\-\_cast<>(...)} respectively with the
following prototypes:

{\scriptsize\begin{verbatim}
template<class T2,
class T1> RefCountPtr<T2> rcp_static_cast(const RefCountPtr<T1>& p1);

template<class T2, class T1>
RefCountPtr<T2> rcp_const_cast(const RefCountPtr<T1>& p1);

template<class T2, class T1>
RefCountPtr<T2> rcp_dynamic_cast(const RefCountPtr<T1>& p1);
\end{verbatim}}

The usage of these conversion template functions looks very similar to
the syntax used for raw C++ pointers.  For example

{\scriptsize\begin{verbatim}
RefCountPtr<const C>   c_ptr   = rcp(new C);
RefCountPtr<const A>   ca_ptr  = c_ptr;                              // Implicit!
RefCountPtr<const C>   cc_ptr1 = rcp_dynamic_cast<const C>(ca_ptr);  // Safe!
RefCountPtr<const C>   cc_ptr2 = rcp_static_cast<const C>(ca_ptr);   // Unsafe!
RefCountPtr<A>         a_ptr   = rcp_const_cast<A>(ca_ptr);          // Cast away const
\end{verbatim}}

Just like the built-in C++ conversion operators, some types of
conversions will not compile.  For example

{\scriptsize\begin{verbatim}
RefCountPtr<C>       c_ptr1 = rcp(new C);
RefCountPtr<A>       a_ptr  = c_ptr;
RefCountPtr<const C> c_ptr2 = rcp_dynamic_cast<const C>(a_ptr);                      // Error!
RefCountPtr<const C> c_ptr3 = rcp_dynamic_cast<const C>(const_cast<const A>(a_ptr)); // Okay!
RefCountPtr<D>  d_ptr  = rcp(new E);
RefCountPtr<E>  e_ptr1 = rcp_dynamic_cast<E>(d_ptr); // Error, D and E are not polymorphic!
RefCountPtr<E>  e_ptr2 = rcp_static_cast<E>(d_ptr);  // Okay, but unchecked!
\end{verbatim}}

This brings us to another commandment.

\begin{commandment}\label{rcp:cmnd:converstion}
Thou shalt only cast between {}\texttt{Ref\-Count\-Ptr<>} objects using the
default copy constructor (for implicit conversions) and the nonmember
template functions {}\texttt{rcp\-\_static\-\_cast<>( ...)},
{}\texttt{rcp\-\_const\-\_cast<>(...)} and
{}\texttt{rcp\-\_dynamic\-\_cast<>(...)}  described above.
\end{commandment}

A developer should never never never try to convert between
{}\texttt{Ref\-Count\-Ptr<>} using the {}\texttt{get()} member function
(which violates Commandment {}\ref{rcp:cmnd:converstion}).  For
example, the following piece of code

{\scriptsize\begin{verbatim}
  {
    ...
    RefCountPtr<A> a_ptr = rcp(new A());
    ...
    RefCountPtr<C> c_ptr = rcp(a_ptr.get());
    ...
  }
\end{verbatim}}

{}\noindent{}will always (unless other steps are taken) result in a
memory deallocation error when these smart pointer objects are deleted
(i.e.~the pointer being deleted twice).

%
\section{Advanced features of {}\texttt{Ref\-Count\-Ptr<>}}
%

In this section we describe some advanced features of
{}\texttt{Ref\-Count\-Ptr<>} that most developers will never need to
use and therefore may never need to know about.  However, there are
certain challenging use cases where this functionality is critical to
the success of {}\texttt{Ref\-Count\-Ptr<>} and therefore are
discussed here.

%
\subsection{Accessing reference-count information}
%

In most situations, clients do not need to know anything about
reference counts but there are use cases where it is helpful if
clients can query reference-count information.  For this purpose,
{}\texttt{Ref\-Count\-Ptr<>} contains the member functions

{\scriptsize\begin{verbatim}
template<class T> int  RefCountPtr<T>::count() const {...}
template<class T> bool RefCountPtr<T>::has_ownership() const {...}
\end{verbatim}}

{}\noindent{}which can be used to ascertain the status of a
reference-counted object.  Specifically, if {}\texttt{ptr} is a
{}\texttt{Ref\-Count\-Ptr<>} object and {}\texttt{(ptr.count()==1 \&\&
ptr.has\_ownership())==true} then the client that owns {}\texttt{ptr}
(where {}\texttt{ptr} is perhaps as a private data member) can infer
that it is the only client that has a reference to the underlying
reference-counted object. Therefore, such a client is free to change
the object without worry that such a change will affect another
client.

This type of information must be used very carefully but can be very
helpful in simplifying the logic in certain types of use cases. If
this type of logic is to be used then it must be documented very
clearly since it is easy to falsely show a reference count higher than
the number of clients that really own {}\texttt{Ref\-Count\-Ptr<>} objects
to the reference-counted object.

%
\subsection{Associating extra data with a {}\texttt{Ref\-Count\-Ptr<>} object}
\label{rcp:sec:extra-data}
%

There are situations when a single {}\texttt{Ref\-Count\-Ptr<>} object
is insufficient to completely handle the dynamic memory for a
reference-counted object.  This can occur, for instance, when trying
to use classes that were written without dynamic memory allocation in
mind.  To demonstrate the issues involved, assume that the following
classes are to be reused in a dynamic environment where the memory
management is to be handled using {}\texttt{Ref\-Count\-Ptr<>}
objects.

{\scriptsize\begin{verbatim}
class Base {
public:
  virtual ~Base() {}
  ...
};

class Derived1 : public Base {
  ...
};

class Utility {
  ...
};

class Derived2 : public Base {
public:
  Derived2(const Utility& u) : u_(u) {}
  ...
private:
  const Utility& u_;
};
\end{verbatim}}

{}\noindent{}Above, every {}\texttt{Derived2} object must have an
associated {}\texttt{Utility} object.  The class {}\texttt{Derived2}
assumes that an object of type {}\texttt{Utility} is allocated
(presumably on the stack or as a global variable) by the client before
it is used to construct a type {}\texttt{Derived2} object.  For
example, the designers of these classes may have expected that class
objects would be used in a flat program such as:

{\scriptsize\begin{verbatim}
void f(Base* b);

int main() {
  ...
  if(...) {
    Derived1 d1;
    f(d1);
  }
  else {
    ...
    Utility u;
    Derived2(u) d2;
    f(d2);
  }
  ...
  return 0;
}
\end{verbatim}}

{}\noindent{}In the above types of programs the flow of logic is
linear from beginning to end and it is easy to allocate all of the
objects on the stack.

Now suppose that we wish to create a factory function

{\scriptsize\begin{verbatim}
RefCountPtr<Base> createBase(bool someFlag);
\end{verbatim}}

{}\indent{}(see the ``Abstract Factory'' design pattern in
{}\cite{ref:gama_et_al_1995}) that dynamically allocates a
{}\texttt{Base} object (of concrete type {}\texttt{Derived1} or
{}\texttt{Derived2}) given a runtime flag and return it using a
{}\texttt{RefCountPtr<Base>} object.  Now let us consider the
following naive and incorrect first implementation of this function.

{\scriptsize\begin{verbatim}
RefCountPtr<Base> createBase(bool someFlag)
{
  if(someFlag) {
    return rcp(new Derived1);
  }
  else {
    Utility u;
    return rcp(new Derived2(u));
  }
}
\end{verbatim}

}{}\noindent{}After some thought, the problem with the above function
should be obvious.  As soon as the block of code where {}\texttt{u} is
allocated (on the stack) is exited, the created {}\texttt{u} object
(and therefore also the {}\texttt{Derived2(u)} object) becomes invalid
before the function even returns.  A second naive and incorrect
implementation of this function might be

{\scriptsize\begin{verbatim}
RefCountPtr<Base> createBase(bool someFlag)
{
  if(someFlag) {
    return rcp(new Derived1);
  }
  else {
    Utility *u = new Utility;
    return rcp(new Derived2(*u));
  }
}
\end{verbatim}}

{}\noindent{} The above implementation will insure that the
{}\texttt{Utility} object created in this function will remain valid
and therefore the {}\texttt{Derived2} object returned from the
function will remain valid but a memory leak will have been created.
If the function {}\texttt{create\-Base(...)}  is only called a few
times in a program then this memory leak will most likely be harmless
but if this function is called too many times, memory exhaustion will
occur.

The solution to this type of problem is to attach a
{}\texttt{RefCountPtr<Utility>} object to the created
{}\texttt{RefCountPtr<Derived2>} object so that when all of the
{}\texttt{RefCountPtr<Derived2>} objects pointed to this object are
deleted (along with its underlying {}\texttt{Derived2} object) the
associated {}\texttt{Utility} object will also be deleted.  The way to
do this is to use the nonmember function

{\scriptsize\begin{verbatim}
template<class T1, class T2>
void set_extra_data(
  const T1 &extra_data, const std::string& name
  ,RefCountPtr<T2> *p, bool force_unique = true
  );
\end{verbatim}}

The function {}\texttt{set\-\_extra\-\_data(...)} adds a piece of
extra data of any type\footnote{Note: the type {}\texttt{T1} used for
extra data must support value semantics (i.e.~have default and copy
constructors defined)} to an existing {}\texttt{RefCountPtr<T2>}
object.  A string \texttt{name} is associated with each piece of extra
data to disambiguate from other pieces of extra data with the same
type and to facilitate later retrieval.  The boolean argument
\texttt{force\-\_unique} (which has a default value of \texttt{true})
is used to allow or disallow overwriting existing data of the same
type and name.  A piece of extra data is keyed both on the type of the
extra data and its string name and access is granted by using one of
the following nonmember functions.

{\scriptsize\begin{verbatim}
template<class T1, class T2>
T1& get_extra_data( RefCountPtr<T2>& p, const std::string& name );

template<class T1, class T2>
const T1& get_extra_data( const RefCountPtr<T2>& p, const std::string& name );
\end{verbatim}}

{}\noindent{} See the Teuchos Doxygen documentation for more details
on these functions.

In many cases, a client does not care to retrieve a set piece of extra
data and therefore the string \texttt{name} really only needs to
disambiguate the extra data being added from other pieces of extra
data that may potentially already be added (perhaps by another unknown
client).  This allows multiple, potentially unrelated, clients to tack
on extra data to a {}\texttt{Ref\-Count\-Ptr<>} object and retrieve
that extra data without stomping on each other.  It is important to
remember that in order for a client to retrieve a piece of extra data
that was added to a {}\texttt{Ref\-Count\-Ptr<>} object, the client
must know the exact type (i.e.~not a base type) of the extra data and
its string name that where used with the call to the
{}\texttt{set\-\_extra\-\_data()} function that added this data.

New let us see how to use the function {}\texttt{set\-\_extra\-\_data(...)}
to correctly implement {}\texttt{create\-Base(...)}

{\scriptsize\begin{verbatim}
RefCountPtr<Base> createBase(bool someFlag)
{
  if(someFlag) {
    return rcp(new Derived1);
  }
  else {
    RefCountPtr<Utility> u = rcp(new Utility);
    RefCountPtr<Base> d = rcp(new Derived2(*u));
    set_extra_data( u, "createBase::u", &d );
    return d;
  }
}
\end{verbatim}}

{}\noindent{}Now when the last {}\texttt{Ref\-Count\-Ptr<>} object to the
dynamically allocated {}\texttt{Derived2} object is removed and the
{}\texttt{Derived2} object is deleted, the associated dynamically
allocated {}\texttt{Utility} object will also be deleted.  Note that
extra data associated with a {}\texttt{Ref\-Count\-Ptr<>} object is only
deleted \underline{after} the reference-counted object is deleted.
This is the correct behavior in use cases like described above where
the destructor for {}\texttt{Derived2} may need to call functions on
its aggregate {}\texttt{Utility} object and therefore the utility
object must be valid at this point.  Added pieces of extra data are
destroyed in a first-in last-out manner (just as is done on the
stack).

Note that the ability to add extra data to a
{}\texttt{Ref\-Count\-Ptr<>} object is an advanced feature that that
most programmers may never need to use.  However, when backed into a
corner, this functionality can be critical to a successful
implementation.

The reason that {}\texttt{get\-\_extra\-\_data()} template functions
are nonmember functions is that the C++ standard does not allow the
explicit specification of a template argument in a member template
function without the use of a verbose, nonintuitive syntax
{}\cite[C.13.6]{ref:stroustrup_2000}.

Even though {}\texttt{set\-\_extra\-\_data()} can deduce the template
type from the input argument, there may be cases where a client may
want to be specific about what type is being used to store extra data
such as in the following case

{\scriptsize\begin{verbatim}
set_extra_data<float>( 1e+5, "some-number", &a_ptr );
\end{verbatim}}

{}\noindent{}In the above example, if the type of the extra data
{}\texttt{float} had not been explicitly specified, the compiler would
have used {}\texttt{double} as dictated by the standard.  However, the
same result could have been achieved with

{\scriptsize\begin{verbatim}
set_extra_data( static_cast<float>(1e+5), "some-number", &a_ptr );
\end{verbatim}}

{}\noindent{}which is more verbose.  It is up to the developer to
decide what form is more desirable.

%
\subsection{Customized deallocation through templated {}\texttt{Dealloc} policy objects}
\label{rcp:sec:dealloc}
%

The most common use of {}\texttt{Ref\-Count\-Ptr<>} is to manage the
lifetime of objects allocated using operator {}\texttt{new} and
deallocated using operator {}\texttt{delete}.  For these use cases,
the built-in behavior in {}\texttt{Ref\-Count\-Ptr<>} does exactly the
right thing.  However, there are situations when one can not simply
call {}\texttt{delete} to deallocate an object.  Some examples of
these situations include:
%
\begin{enumerate}
\item
When reference counts for objects are managed by requiring clients to
explicitly call increment and decrement functions.  This situation
occues when using CORBA {}\cite{ref:corba} and COM {}\cite{ref:com}
for instance.  Such an approach is also presented in {}\cite[Item
29]{ref:meyers_1996} in the subsection ``A Reference-Counting Base
Class''.  In these protocols, deallocation occurs automatically behind
the scenes when this other reference count goes to zero and does not
occur through an explicit call to operator {}\texttt{delete} as with
the default behavior for {}\texttt{Ref\-Count\-Ptr<>'}.
\item
When objects are managed by certain types of object databases.  In
some object databases, an object that is grabbed from the database must
be explicitly returned to the database in order to allow proper object
deletion to take place later.
\item
A different reference-counted pointer class is used to initially get
access to the managed object.  For example, suppose some piece of peer
software works with {}\texttt{boost::shared\_ptr<>} (see
{}\cite{ref:boost}) referenced-counted objects while the resident
software works with {}\texttt{Ref\-Count\-Ptr<>} objects.  It then
becomes critical no object is deleted until all the clients using
either of these smart pointer types remove their references
(i.e.~destroy their smart pointers) to this underlying object.
\end{enumerate}
%
There are also other situations where one can not simply assume that
calling operator {}\texttt{delete} is used to release an object.  The
bottom line is that in order to be general, one must allow arbitrary
policies to be used to deallocate an object after clients are finished
using the object.

{}\texttt{Ref\-Count\-Ptr<>} supports arbitrary deallocation polices
using the concept of a template deallocator object through the
following version of the nonmember function {}\texttt{rcp(...)}.

{\scriptsize\begin{verbatim}
template<class T, class Dealloc_T>
RefCountPtr<T> rcp( T* p, Dealloc_T dealloc, bool owns_mem );
\end{verbatim}}

The implementation of the simpler version of {}\texttt{rcp(...)} first
described in Section {}\ref{rcp:sec:init-rcp-objects} uses the default
deallocator class

{\scriptsize\begin{verbatim}
template<class T>
class DeallocDelete
{
public:
  typedef T ptr_t;
  void free( T* ptr ) { if(ptr) delete ptr; }
};
\end{verbatim}}

All deallocator objects must support the typedef member
{}\texttt{ptr\_t} and function member {}\texttt{free()}.  The concept
of a template policy interface (also called a function object
\cite[Section 18.4]{ref:stroustrup_2000}) should be familiar to
semi-advanced users of the STL (part of the standard C++ library).

To demonstrate the use of a deallocator object, let us assume that we
must wrap objects of type {}\texttt{A} managed by the the following
object database

{\scriptsize\begin{verbatim}
class ObjectADB {
  ...
  A& get(int id);
  void release(int id);
  ...
};
\end{verbatim}}

In the above object database, objects are accessed and released using
an id.  How this id is specified and determined is not important here.
Now let us suppose that we want to define an abstract factory that
returns objects of type {}\texttt{A} wrapped in
{}\texttt{RefCountPtr<A>} objects using a database of type
{}\texttt{ObjectADB} shown above.  For this abstract factory, objects
of type {}\texttt{A} will be allocated from a list of ids given to the
factory.  The outline of this abstract factory subclass looks like the
following

{\scriptsize\begin{verbatim}
class ObjectADBFactory : public AbstractFactory<A> {
public:
  ObjectADBFactory( ObjectADB *db, int num_ids, const int ids[] ) : db_(db), ids_(ids,ids+num_ids) {}
  RefCountPtr<A> create() { ... };  // Overridden from AbstractFactory
private:
  ObjectADB *db_;
  std::vector ids_;
};
\end{verbatim}}

The above abstract factory subclass {}\texttt{ObjectADBFactory}
inherits from a generic {}\texttt{AbstractFactory} base class that
defines a pure virtual method {}\texttt{create()}.  In order to
implement the override of the {}\texttt{create()} method, a
deallocator class must be defined and used.  For this purpose we
define

{\scriptsize\begin{verbatim}
template<class A>
class DeallocObjectADB
{
public:
  DeallocObjectADB( ObjectjADB* db, int id ) : db_(db), id_(id) {}
  typedef A ptr_t;
  void free( T* ptr ) { db_->release(id); }
private:
  ObjectjADB* db_;
  int id_;
  DeallocObjectADB(); // not defined and not to be called!
};
\end{verbatim}}

{}\noindent{}Above, the default constructor is declared private and
has no implementation to avoid accidental default construction (see
Item 27 in {}\cite{ref:meyers_1994}).

Now we can define the implementation of the {}\texttt{create()} function
override below.

{\scriptsize\begin{verbatim}
RefCountPtr<A> ObjectADBFactory::create()
{
  TEST_FOR_EXCEPTION( ids_.size()==0, std::runtime_error, "No ids are left!" );
  const int id = ids_.pop();
  return rcp(&db_->get(id),DeallocObjectADB(db_,id),true);
}
\end{verbatim}}

The following program shows the use of the above factory class

{\scriptsize\begin{verbatim}
int main()
{
  // Create the object database and populate it (and save the ids)
  ObjectADB db;
  std::vector ids;
  ...
  // Create the abstract factory object
  ObjectADBFactory  ftcy(db,ids.size(),&ids[0])
  // Create some A objects and use them
  RefCountPtr<A> a_ptr1 = fcty.create();
  ...
  return 0;
}
\end{verbatim}}

In the above program, all of the objects of type {}\texttt{A} are
created and removed seamlessly without the client code that interacts
with {}\texttt{Ref\-Count\-Ptr<>} and {}\texttt{AbstractFactory<>}
knowing anything about what is going on under the hood.

%
\section{Guidelines for the usage of {}\texttt{Ref\-Count\-Ptr<>} and other C++ data types}
%

After the basics of {}\texttt{Ref\-Count\-Ptr<>} are understood its
use is quite straightforward.  Furthermore,this class can be
instantiated with any built-in or user-defined concrete or abstract
data type, so one may ask why {}\texttt{Ref\-Count\-Ptr<>} should not
be used for all objects.  There is nothing stopping a developer from
using {}\texttt{Ref\-Count\-Ptr<>} in this way but to do so would be
overkill and counterproductive. The most compelling situation where
{}\texttt{Ref\-Count\-Ptr<>} should be used is when a client needs to
maintain a private data member to another object which is of an
abstract type or is an object that is shared with other clients.  As
described in the beginner's guide
{}\cite{ref:RefCountPtrBeginnersGuide} this is an example of a
{}\textit{non-trnasitive} relationship and
{}\texttt{Ref\-Count\-Ptr<>} should be used for all
{}\textit{non-trnasitive} relationships.  Note that
{}\texttt{Ref\-Count\-Ptr<>} usually should not be used for built-in
data types such as {}\texttt{int} or {}\texttt{double} unless there is
a desire to allow one client to change the value and then have this
value automatically updated in other clients (which is rarely a good
idea to allow in general).

Note that the consistent use of {}\texttt{Ref\-Count\-Ptr<>} makes it
possible to skip writing explicit destructors for most classes since
the compiler-generated destructor does exactly the right thing.

%
\section{Idioms for passing objects to and from functions}
\label{rcp:sec:passing-args}
%

This section focuses on how to best write function prototypes using
{}\texttt{Ref\-Count\-Ptr<>}.  Understanding how to use
{}\texttt{Ref\-Count\-Ptr<>} well also requires understanding how to
use more standard concrete and abstract data types through raw C++
references and pointers.

Choosing how to write function prototypes for passing arguments to and
from functions in languages like C and Fortran 77 (exception F77 does
not have a concept of function prototypes) is simple since there are
few choices to make.  In C you can either pass an object by value or
by reference (i.e.~through a pointer).  In Fortran 77 the choice is
even easier (there is no choice) since every variable is passed by
reference.  Passing arrays of objects to and from functions in these
languages is slightly more complicated, but not by much.

On the other hand, C++ adds the concepts of {}\texttt{const} and
references (i.e.~{}\texttt{type\& obj}) for passing arguments in
addition to C's concepts of pointers (i.e.~{}\texttt{type* obj}) and
pass-by-value (i.e.~{}\texttt{type obj}).  This makes deciding how to
pass arguments in C++ much more complicated than in other languages.
These features (which where added to C++ for very good technical
reasons), when guided by commonsense idioms, can make it more-or-less
clear how arguments should be passed to and from functions in C++
based on their purpose and can serve to clearly documument the purpose
in many cases without having to write much extra documentation.  This
discussion presents some of these idioms and how
{}\texttt{Ref\-Count\-Ptr<>} fits in.

In many respects C++ is a highly expressive programming language where
if used carefully, in many cases, the programmer's intent for an
interface or a function prototype can be discerned just by looking at
the C++ declaration.  For example, if a reference or pointer argument
passed into a function has the {}\texttt{const} qualifier, then a
developer can rightly assume that the argument will not be changed
within the function.  The language tries to help enforce
{}\texttt{const} but, in the end, it is up to the implementer of the
function to honor the promise inherent in the {}\texttt{const}
modifier in the function prototype (for a more detailed discussion of
{}\texttt{const} in C++ see {}\cite[Section
10.2.7.1]{ref:stroustrup_2000}).  In other cases, the exact intent of
an argument passed to a function can not be discerned just be looking
at the function prototype.  For example, if an array argument is
specified as {}\texttt{int a[]} it is not clear if this array is
required (i.e.~{}\texttt{a!=NULL} required) or is optional
(i.e.~{}\texttt{a==NULL} allowed).  C++ is just not expressive enough
to completely document every use case for an argument.

In addition to the idioms mentioned above, there are other C++ idioms
that can be used so that the general nature of arguments passed to and
from a function can be largely discerned just by looking at the C++
prototype for the function.  The {}\texttt{const} qualifier just
mentioned above is one of the language features which is
self-documenting but there are others.  Another example is passing
arrays to and from functions, for example, by declaring an integer
array argument as {}\texttt{int arg[]} instead of {}\texttt{int* arg}.
When declaring an argument as {}\texttt{int arg[]} there is no
question that an array is being passed but in the case of
{}\texttt{int* arg} it is not clear if the address of the first
element of an array is being passed or just the address to a single
object.  Appendix D in the beginner's guide
{}\cite{ref:RefCountPtrBeginnersGuide} shows recommendations for how
to pass objects to and from C++ functions in a way that is as
self-documenting as possible.

The reasons why the declarations are the way they are in Appendix D in
{}\cite{ref:RefCountPtrBeginnersGuide} can be found in the references
{}\cite{ref:stroustrup_2000} and {}\cite{ref:meyers_1994}.
Specifically, see the sections ???.  The discussion below describes
each of the use cases shown in Appendix D in
{}\cite{ref:RefCountPtrBeginnersGuide} and how
{}\texttt{Ref\-Count\-Ptr<>} fits in.

As a basic rule of thumb, {}\texttt{Ref\-Count\-Ptr<>} objects should
be passed around as though they where raw C++ pointers to the
underlying objects.  The main difference between passing
{}\texttt{Ref\-Count\-Ptr<>} and and raw C++ pointers, however, is
that it usually a good idea to pass {}\texttt{Ref\-Count\-Ptr<>}
objects by {}\texttt{const} reference instead of by value since
passing {}\texttt{Ref\-Count\-Ptr<>} objects by value results in
unnecessary calls to copy constructors and destructors that, if
nothing else, make stepping through code in debugger more difficult.

%
\subsection{Detailed discussion/justification for idioms}
%

ToDo: Put in a lot more example code.

If the reader already accepts and is confortable with the idioms for
function prototypes shown in {}\cite[Appendix
D]{ref:RefCountPtrBeginnersGuide} and overviewed in the previous
section, then this section can be skipped until later.  However,
readers that do not agree or are curious why the idioms in
{}\cite[Appendix D]{ref:RefCountPtrBeginnersGuide} are emphasised in
this document should read the following section and them make up their
own minds.  This material is not fundumental to the design of
{}\texttt{Ref\-Count\-Ptr<>} but instead is more basic advise about
good C++ practicies along with experienced guidelines for the use of
{}\texttt{Ref\-Count\-Ptr<>}.

We now explain the reasoning behind the idiom recommendations given in
Appendix D in {}\cite{ref:RefCountPtrBeginnersGuide} for how to pass
objects to and from C++ functions and how {}\texttt{Ref\-Count\-Ptr<>}
fits in with these idioms.  First it should be stated that many
different code projects do not follow these idioms.  Usually this is
because the developers where unaware of these idioms and why they are
useful to follow.  Even if a developer is familiar with these idioms
for raw data types, the may not be as familiar with how
{}\texttt{Ref\-Count\-Ptr<>} fits in.

As stated in the beginner's guide
{}\cite{ref:RefCountPtrBeginnersGuide}, it is recommended to use raw
object references for passing transitive arguments (i.e.~those that
are only accessed during a function call) to a function.  Here, the
term reference is used in a more abstract sense where a reference can
be implemented as a C++ reference such as {}\texttt{const A \&a} or as
a C++ pointer such as {}\texttt{A *a}.  On the other hand, it is
recommended to pass {}\texttt{Ref\-Count\-Ptr<>} wrapped object
references (i.e.~{}\texttt{const Ref\-Count\-Ptr<A> \&a}) for any
non-transitive arguments (i.e.~those where a ``memory'' of the object
in the server implementing the function persists after the function
returns).

Passing {}\texttt{Ref\-Count\-Ptr<>} wrapped object references to and
from functions will differ slightly from how raw object references are
passed.  As a rule of thumb, when thinking about how to use
{}\texttt{Ref\-Count\-Ptr<>} to pass object references to and from
functions, one should think of a {}\texttt{Ref\-Count\-Ptr<>} object
as though it was really just a raw pointer.  The main difference is
that while raw pointers to single objects are usually passed by value
(i.e.~{}\texttt{A *a}) to a function (which is very efficient) a
{}\texttt{Ref\-Count\-Ptr<>} wrapped single object on the other hand
should usually be passed by {}\texttt{const} reference
(i.e.~{}\texttt{const Ref\-Count\-Ptr<A> \&a}).  It is important to
pass single {}\texttt{Ref\-Count\-Ptr<>} object by {}\texttt{const}
reference and not by non-{}\texttt{const} reference
(i.e.~{}\texttt{Ref\-Count\-Ptr<A> \&a}) since the compiler will only
perform automatic conversions for {}\texttt{const} references.
However, passing {}\texttt{Ref\-Count\-Ptr<>} objects by value is
harmless but passing {}\texttt{Ref\-Count\-Ptr<>} objects by
{}\texttt{const} reference eliminates unnecessary (while quite
trivial) constructor and destructor function calls.  In any case,
passing {}\texttt{Ref\-Count\-Ptr<>} objects by {}\texttt{const}
reference as apposed to by value simplifies stepping through code in a
debugger and should be preferred if only for this reason.  On the
other hand, when passing arrays of objects wrapped within
{}\texttt{Ref\-Count\-Ptr<>} objects, it is recommended to pass an
array of {}\texttt{Ref\-Count\-Ptr<>} objects (i.e.~{}\texttt{const
Ref\-Count\-Ptr<A> a[]}) and not an array of
{}\texttt{Ref\-Count\-Ptr<>} references.  This is discussed below.

Before getting into great detail about these recommendations and
exactly how {}\texttt{Ref\-Count\-Ptr<>} fits in, consider the
following example function prototype.

{\scriptsize\begin{verbatim}
void foo1(
  const int              size_x
  ,const double          x[]
  ,const int             size_a
  ,const A*              a[]
  ,const int             size_y
  ,double                y[]
  ,double                *z
  );
\end{verbatim}}

By comparing the argument declarations (which use good argument names)
in the above function to the recommendations in {}\cite[Appendix
D]{ref:RefCountPtrBeginnersGuide}, and without looking at any extra
documentation (if it even exists), we already know a lot about the
arguments being passed into and out of this function.  First, it is
clear that {}\texttt{x} is an array of non-mutable {}\texttt{double}
objects with an presumed length of {}\texttt{size\_x}.  Similarly, we
know that {}\texttt{y} is an array of mutable {}\texttt{double}
objects that may be set or modified.  We also know that {}\texttt{z}
is a single mutable {}\texttt{double} object that may be changed
inside of the function.  What the above C++ declarations, however, do
not tell us is whether the arguments {}\texttt{x}, {}\texttt{a},
{}\texttt{y} and {}\texttt{z} are optional (i.e.~a {}\texttt{NULL}
pointer can be passed in for them) or are required (i.e.~a
non-{}\texttt{NULL} pointer can be passed in for them).  We also do
not know if the mutable objects passed in {}\texttt{y} and
{}\texttt{z} are only output arguments (i.e.~the state of the object
before the function call is not significant) or are input/output
arguments (i.e.~the state of the object before the function call is
significant).  This type of information must be specified in
developer-supplied documentation.  Note that while a required single
non-mutable object such as {}\texttt{z} could be passed using a
non-{}\texttt{const} reference, using a non-{}\texttt{const} reference
for such an argument is usually not advised and this issue is
addressed a little later.

Note that the above function {}\texttt{foo(...)} involved all
transitive (i.e.~only accessed with the function) arguments which is
often the case.  Now lets consider the use of non-transitive argument
types as supported using {}\texttt{Ref\-Count\-Ptr<>}.

{\scriptsize\begin{verbatim}
class SomeClass {
public:
  ...
  void foo2(
    const int                                 size_x
    ,double                                   x[]
    ,const RefCountPtr<std::vector<double> >  &y
    )
  {
    y_ = y;
    ...
  }
  ..
private:
  RefCountPtr<std::vector<double> >  y_;
};
\end{verbatim}}

{}\noindent{}The above example function
{}\texttt{SomeClass\-::foo2(...)} is an example where a non-transitive
argument, {}\texttt{y}, is passed into a function and is then
``remembered''.  In this case, the {}\texttt{Ref\-Count\-Ptr<>}
argument {}\texttt{y} is delivering a
{}\texttt{std\-::vector<\-double>} object to be used as private data
for the class object.

%
\subsubsection{Passing non-mutable objects}
%

Non-mutable transitive objects (i.e.~objects that are passed into a
function, are not modified and no memory of these objects is
maintained) can be passed by value or by reference (i.e.~through a C++
pointer or a C++ reference) and as either single objects or as an
array of objects.  When non-mutable objects are passed to a function
using a C++ pointer or a C++ reference, the declaration should always
use the {}\texttt{const} modifier.  Failure to use the
{}\texttt{const} modifier in these cases goes against the design of
the C++ language and is incompatible with the standard C++ library.
The use of the {}\texttt{const} modifier both serves as documentation
and helps the implementor of the function avoid accidentally changing
the object.  However, when a small concrete object of type
{}\texttt{S} (where {}\texttt{S} is a {}\texttt{double} or
{}\texttt{int} for instance) is being passed, it may also be passed by
value (with or without the {}\texttt{const} modifier as shown in
{}\cite[Appendix D]{ref:RefCountPtrBeginnersGuide}).  When a concrete
object is passed by value, the {}\texttt{const} modifier does not in
any way affect the client code that calls the function and the
client's copy of the argument is guaranteed not to be changed
regardless of whether the {}\texttt{const} modifier is used or not.
On the other hand, the {}\texttt{const} modifier on a C++ pointer or
C++ reference does not really guarantee that the object will not be
changed.  The {}\texttt{const} modifier used with pass-by-value only
restricts the modification of the copied argument in the function
implementation.  Some see the use of {}\texttt{const} pass-by-value as
a means to help function implementors from making mistakes while
others see this as a unnecessary restriction on the freedom in the
implementation of the function.

When a non-mutable input argument is required, it should be passed by
value (i.e.~{}\texttt{S s}) or by {}\texttt{const} C++ reference
(i.e.~{}\texttt{const S \&s}).  However, when a non-mutable input
argument is optional, it must be passed using a {}\texttt{const} C++
pointer (i.e.~{}\texttt{const S *s}) since this pointer is be allowed
to be {}\texttt{NULL} for when the argument is not specified.

Arrays of non-mutable objects should either be passed as an array of
objects (i.e.~in the case where the object's type is a small concrete
data type such as {}\texttt{double}) such as {}\texttt{S s[]} or
{}\texttt{const S s[]}, or as an array of {}\texttt{const} C++
pointers (i.e.~in the case where the type is a large object or the
type has reference semantics) such as {}\texttt{const S* s[]}.  Note
that the syntax {}\texttt{const S s[]} (or {}\texttt{const A* a[]})
for passing an array of objects is to be preferred over the C++
pointer syntax {}\texttt{const S *s} when passing an array of objects
or {}\texttt{const A** a} when passing an array of pointers.  This is
because that even though the raw C++ pointer syntax {}\texttt{const S
*s} is perfectly legal and proper C++ usage, this form does not allow
one to tell if this pointer is supposed to point to a single object or
to an array of objects and therefore the syntax {}\texttt{const S s[]}
is to be preferred.  Also note that the syntax {}\texttt{const A* a[]}
allows individual components of {}\texttt{a[i]} to be optional
(i.e.~{}\texttt{a[i]==NULL} allowed) or required
(i.e.~{}\texttt{a[i]!=NULL} required) in addition to the entire array
being optional (i.e.~{}\texttt{a==NULL} allowed) or required
(i.e.~{}\texttt{a!=NULL} required).  In this case the preconditions
for these types of array of pointer arguments must be spelled out in
supplemental documentation.

Note that while it is possible to pass objects as an array of object
references such as {}\texttt{const S\& s[]}, it is generally very
difficult to initialize the references right when the array is first
created (as is required by the standard).  Therefore it is not
recommended to passed objects as an array of references but instead to
use an array of pointers as shown in {}\cite[Appendix
D]{ref:RefCountPtrBeginnersGuide}

Also note that unlike in the case of passing single non-mutable
objects, that there is no a convenient way to help differentiate
whether an array of objects being passed to a function is optional or
not.  When passing arrays of objects, it must be documented whether
the arrays are required or optional.

Non-mutable non-transitive objects (i.e.~objects that are passed into
a function, are not modified but some memory of these objects is
maintained after the function call) should be passed through
{}\texttt{Ref\-Count\-Ptr<>} wrapped objects.  Single object
references of this type should be passed as {}\texttt{const
Ref\-Count\-Ptr<const A> \&a}.  Note the presences of the
{}\texttt{const} templated type {}\texttt{<const A>} which states that
clients may not modify the underlying wrapped object (see
{}\cite[Appendix B]{ref:RefCountPtrBeginnersGuide} and Section
{}\ref{rcp:sec:init-rcp-objects} for an explanation of how to express
combinations of {}\texttt{const} or non-{}\texttt{const} pointers to
{}\texttt{const} or non-{}\texttt{const} objects using
{}\texttt{Ref\-Count\-Ptr<>}).  As stated earlier, the
{}\texttt{Ref\-Count\-Ptr<>} object itself should be passed as a
{}\texttt{const} reference to avoid unnecessary constructor and
destructor calls.  Note that unlike passing single raw
{}\texttt{const} object reference to a function, there is no easy way
at compile time to differentiate between a required and an optional
argument when passing a {}\texttt{Ref\-Count\-Ptr<>} wrapped object.
For the argument {}\texttt{const Ref\-Count\-Ptr<const A> \&a}, if the
underlying object is required, then the precondition
{}\texttt{a.get()!=NULL} should be documented.  Likewise, if the
object is optional, then it should be documented that
{}\texttt{a.get()==NULL} is allowed.

Passing an array of non-mutable non-transitive objects wrapped in
{}\texttt{Ref\-Count\-Ptr<>} objects is a little different from arrays
for raw C++ objects or object references.  While it is recommended to
pass single object references using {}\texttt{const}
{}\texttt{Ref\-Count\-Ptr<>} C++ references, an array of
{}\texttt{Ref\-Count\-Ptr<>} objects should be passed as an array of
{}\texttt{Ref\-Count\-Ptr<>} objects as {}\texttt{const
Ref\-Count\-Ptr<const A> a[]} rather than trying to pass an array of
C++ references to {}\texttt{Ref\-Count\-Ptr<>} objects such as
{}\texttt{const Ref\-Count\-Ptr<const A>\& a[]}.  As stated earlier,
it is extremely difficult to work with arrays of references.  Note
that just as is the case when passing an array of raw object
references, there is no good way to specify at compile time whether
the entire array {}\texttt{Ref\-Count\-Ptr<>} objects is required or
is optional and this instead must be explicitly documented.  Note also
that just as in the case for passing arrays of raw object references,
the declaration {}\texttt{const Ref\-Count\-Ptr<const A> a[]} allows
individual components of {}\texttt{a[i]} to be optional
(i.e.~{}\texttt{a[i].get()==NULL} allowed) or required
(i.e.~{}\texttt{a[i].get()!=NULL} required) in addition to the entire
array being optional (i.e.~{}\texttt{a==NULL} allowed) or required
(i.e.~{}\texttt{a!=NULL} required).  The specific preconditions on
such array arguments must be documented.

%
\subsubsection{Passing mutable objects}
%

Many of the issues involved with passing mutable objects (i.e.~client
maintained objects that are modified within a function) are the same as
for passing non-mutable objects such as described above but there are
some differences.  First off, objects that are to have their state
changed within a function must be passed by reference (i.e.~C++
pointer or C++ reference) and can not be passed by value.  In the case
of mutable arguments, non-{}\texttt{const} instead of {}\texttt{const}
C++ pointers are used to pass raw object references.  Likewise,
{}\texttt{Ref\-Count\-Ptr<A>} objects are used instead of
{}\texttt{Ref\-Count\-Ptr<const A>} objects.  Other than that, all of
the issues are exactly the same as in the case of passing non-mutable
objects.

One apparent area of controversy is the issue of whether
non-\texttt{const} reference or non-\texttt{const} pointer arguments
should be used for passing mutable objects to a function.  If the
argument is optional, then the choice is clear and the argument should
be declared as a non-\texttt{const} pointer.  However, if the argument
is required then some would say that a non-\texttt{const} reference
should always be used.  Interested readers should note, however, that
references were added to C++ primarily to support operator overloading
(see {}\cite[Section 3.7]{ref:design_evol_cpp}) and also for return
values.  If one is writing an overloaded operator function that
modifies one of its arguments (such as the stream insertion and
extraction operators {}\texttt{std::istream\& operator<<(std::istream
\&i, ... )} and {}\texttt{std::ostream\& operator>>(std::ostream \&o,
... )} for instance) then a non-\texttt{const} reference argument must
be used (and note the non-\texttt{const} reference return type as
well).  When the function is not an overloaded operator function then
the choice is not so clear.  Strustroup (the original inventor of C++
{}\cite{ref:design_evol_cpp}) makes a case for using
non-\texttt{const} pointers for mutable arguments in most non-operator
functions in {}\cite[Section 5.5]{ref:stroustrup_2000}.  In short, the
use of non-{}\texttt{const} C++ references over non-{}\texttt{const}
C++ pointer arguments for required mutable objects gives a false sense
of security and in the end does not really guarantee better quality
software.  While that standard states that every C++ reference must be
bound to a valid C++ object this is not guaranteed in any way.  For example,
consider the following program that will segfault when run.

{\scriptsize\begin{verbatim}

#include <iostream>

void f( int& i )
{
  i = 5;
  std::cout << "\ni = "<<i<<std::endl;
}

void g( int* i )
{
  f(*i);
}

int main()
{
  int a  = 2, *b = NULL;
  g(&a);
  g(b);
  return 0;
}

\end{verbatim}}

The above program can be put into a file ({}\texttt{bad\_ref.cpp} for
instance) and be compiled and linked with no errors by any standard
compliant C++ compiler (for instance using g++ as {}\texttt{g++ -o
bad\_ref.exe bad\_ref.cpp}).  However, when this program is run (as
{}\texttt{./bad\_ref.exe} for instance) the output will look something
like:

{\scriptsize\begin{verbatim}

#./bad_ref.exe

i = 5
Segmentation fault (core dumped)

\end{verbatim}}

The reason for the segmentation fault should be quite obvious.  In the
case of the above flat program, a very smart compiler might be able to
produce a warning but in the general case, where the main program and
the functions {}\texttt{f()} and {}\texttt{g()} are compiled in
separate source files, there is no way a C++ compiler can possibly
catch this error at compile time.  This example just proves that the
language does not guarantee that C++ references are bound to valid
objects.

Since non-{}\texttt{const} C++ references don't really give any more
guarantees than non-{}\texttt{const} C++ pointers, the chose of
whether to use non-{}\texttt{const} C++ references over
non-{}\texttt{const} C++ pointers should be made on other grounds.
The primary justification for using non-{}\texttt{const} C++ pointer
arguments for mutable objects is to force a different specification
for non-mutable and mutable objects being passed into a C++ function
that both give visual clues of what objects are going to be changed
and provides some compile-time checking if arguments change from being
non-mutable to mutable.  For example, consider the following program

{\scriptsize\begin{verbatim}

#include <iostream>

void f( const std::vector<int> &i, const std::vector<int> &j, int *k );

int main()
{
  std::vector<int>  i(5,0), j(5,1);
  int k;
  ...
  f( i, j, &k );
  ...
  std::foreach( j.begin(), j.end(), ... );
  ...
  return 0;
}

\end{verbatim}}

In the above program, even if you never even saw the prototype
for the function {}\texttt{f()}, you can assume that the argument
{}\texttt{k} is being modified by the function because of the visual
clue {}\texttt{\&k} in the statement

{\scriptsize\begin{verbatim}
  f( i, j, &k );
\end{verbatim}}

{}\noindent{}while rightly assuming that the arguments {}\texttt{i}
and {}\texttt{j} will not be modified.  Later on in this program,
the statement

{\scriptsize\begin{verbatim}
  std::foreach( j.begin(), j.end(), ... );
\end{verbatim}}

{}\noindent{}uses the array {}\texttt{j}, which is assumed to still be
in its initialized state.  Now suppose that some overly clever
developer decides that he or she can implement the function
{}\texttt{f()} better if the array in {}\texttt{j} is used as
workspace in addition to providing data for the function.  This is a
common practice (especially in well-written Fortran programs).  If
non-\texttt{const} references are the norm for mutable arguments, then
this developer may decide to change the prototype of {}\texttt{f()} to

{\scriptsize\begin{verbatim}
void f( const std::vector<int> &i, std::vector<int> &j, int *k );
\end{verbatim}}

{}\noindent{}which now allows the function to use {}\texttt{j} as
workspace.  Removing a {}\texttt{const} modifier from a function like
this may cause some calls to this function to not compile (which is a
very good thing) but calls like in the above example program will
still compile just fine.  If this change is made to the function
{}\texttt{f()} and the argument {}\texttt{j} is silently changed
without any clue to the code in {}\texttt{main()}, then the behavior
of the rest of the code that follows the call to {}\texttt{f()} may be
silently changed (such as was the case in this function).  This could
be a very hard bug to track down.

On the other hand, if non-\texttt{const} C++ pointer arguments are
always used for mutable objects, then the prototype for {}\texttt{f()}
must be changed to 

{\scriptsize\begin{verbatim}
void f( const std::vector<int> &i, std::vector<int> *j, int *k );
\end{verbatim}}

{}\noindent{}if the argument {}\texttt{j} is to be used a workspace.
In this case, the function call

{\scriptsize\begin{verbatim}
  f( i, j, &k );
\end{verbatim}}

{}\noindent{}would not compile and the developer that maintains this
code would have a great opportunity to change the calling code to cope
with the fact that {}\texttt{j} is no longer a non-mutable argument.
In this case, either a copy of {}\texttt{j} could be created to pass
into {}\texttt{f()} or the logic after the call to {}\texttt{f()}
could no longer assume that {}\texttt{j} is unchanged.

In summary, all things being equal, tracking down runtime errors
associated with using raw non-\texttt{const} C++ pointer arguments is
easier than tracking down runtime errors that result from using
non-\texttt{const} C++ reference arguments and using
non-\texttt{const} C++ pointer arguments give a visual clue as to what
arguments are being changed and what arguments are being left
unmodified.  This is a subjective opinion but is one that is held by
many esteemed C++ experts (including Strustroup himself).

However, while C++ references where {}\underline{not} purposefully
added to the language to support non-\texttt{const} pass-by-reference
in non-operator functions, there are circumstances where using
non-\texttt{const} reference arguments is to be preferred.  An example
might be one where the fact that an argument is going to be changed as
a result of a function call is obvious from the name of the function
or of the nature of the object being passed.  This is the case, for
instance, in the standard C++ function

{\scriptsize\begin{verbatim}
template<class T1, class T2> void swap( T1 &t1, T2 &t2 );
\end{verbatim}}

{}\noindent{}where it is obvious that when called like:

{\scriptsize\begin{verbatim}
void foo3()
{
  int a = 5, b = 6;
  std::swap(a,b);
}
\end{verbatim}}

{}\noindent{}that the objects {}\texttt{a} and {}\texttt{b} will be
modified.

Another example of when it is reasonable to use a non-\texttt{const}
reference is when working with objects such as standard streams
(i.e.~{}\texttt{std::ostream} and {}\texttt{std::istream}) where the
{}\texttt{const} interface is almost worthless.  For example, it is
perfectly reasonable to declare a print function such as

{\scriptsize\begin{verbatim}
class Foo2 {
public:
  ...
  void print(std::ostream &out) const;
  ...
};
\end{verbatim}}

{}\noindent{}where when called as

{\scriptsize\begin{verbatim}
void foo4(const Foo2 &foo2) {
  foo2.print(std::cout);
}
\end{verbatim}}

{}\noindent{}it is obvious that the stream object {}\texttt{std::cout}
will be modifed in the function call.

The general approach advocated here is to prefer non-\texttt{const}
C++ pointer arguments in non-operator functions unless there is a
compelling reason to use a non-\texttt{const} C++ reference instead
such as when there is unlikely to be any confusion whether the
argument is going to be changed and a couple of examples where given
above.  Making the determination of when to use a non-\texttt{const}
C++ reference over a non-\texttt{const} C++ pointer requires
experience, taste and in the end is a subjective decision.

While it was stated earlier that a {}\texttt{Ref\-Count\-Ptr<>} object
should almost always be passed as a {}\texttt{const} reference
(i.e.~{}\texttt{const Ref\-Count\-Ptr<>\&}), there is one use case
where a pointer to a {}\texttt{const} {}\texttt{Ref\-Count\-Ptr<>}
object (i.e.~{}\texttt{Ref\-Count\-Ptr<>*}) should be passed instead.
This occurs when the calling client code is to acquire a
{}\texttt{Ref\-Count\-Ptr<>} to an object held by the server code, or
in other words form non-transitive relation between the returned
object and the client.  When only a single
{}\texttt{Ref\-Count\-Ptr<>} is to be acquired for a single object,
then it usually best to just return a {}\texttt{Ref\-Count\-Ptr<>}
object from the function as its return value such as in the case

{\scriptsize\begin{verbatim}
RefCountPtr<Base> createBase(bool someFlag);
\end{verbatim}}

{}\noindent{}described in Section {}\ref{rcp:sec:extra-data}.
However, when more than one {}\texttt{Ref\-Count\-Ptr<>} object is to
be initialized and given to a client, then one should consider passing
non-{}\texttt{const} pointers to {}\texttt{Ref\-Count\-Ptr<>} objects
to a function (see the function {}\texttt{uninitialize()} in Appendix
{}\ref{rcp:apdx:sep-construct-init}).

%
\section{Summary}
%

This document expanded on the introductory discussion in the
beginner's guide {}\cite{ref:RefCountPtrBeginnersGuide}.  Some of the
more advanced features of {}\texttt{Ref\-Count\-Ptr<>} where described
in more detail.  Suggestions for the usage of
{}\texttt{Ref\-Count\-Ptr<>} and C++ itself was also discussed.  The
Appendices discuss some other topics related to
{}\texttt{Ref\-Count\-Ptr<>} in detail such as the ``separate
construction and initialization'' idiom (Appendix ???), the design of
{}\texttt{Ref\-Count\-Ptr<>} (Appendix ???) and issues related to
multiple inheritance and {}\texttt{Ref\-Count\-Ptr<>} (Appendix ???).
