\documentclass{csri10}

% PACKAGES ---------------------------------------------------------------
\usepackage{amsfonts,amsmath,graphicx,subfigure}
\usepackage{uml}
% ADD YOUR OWN PACKAGES HERE ---------------------------------------------
%\usepackage{someotherpackage}
\usepackage{url}
\usepackage{multicol}
\usepackage{paralist}

% DEFINITIONS ------------------------------------------------------------
% ADD YOUR OWN DEFINITIONS HERE ------------------------------------------
% BE SURE TO PREFACE LABEL WITH YOUR OWN INITIALS (SSS in this example) --
\newcommand{\SSSnorm}[1]{\left\Vert#1\right\Vert}
\newcommand{\SSSabs}[1]{\left\vert#1\right\vert}

% This controls the table-of-contents entry in the proceedings. Edit it
% to include your article title followed by the authors' names, as shown.
\addcontentsline{toc}{chapter}{Optika: A GUI Framework for Parameterized Applications\\
{\em Kurtis Nusbaum Student and Dr.\ Mike Heroux Mentor}}

\pagestyle{myheadings}

\thispagestyle{plain}

% This gives the running head. Usually you list a shortened version of
% your article title (unless it's already very short) along with
% the author's names, as shown.
\markboth{Optika GUI Framework}{Kurtis Nusbaum Student and Dr.\ Mike Heroux Mentor}

% Put your article title in here
\title{Optika: A GUI Framework for Parameterized Applications}

% List each author, their affiliation, and their e-mail address, as shown.
\author{Kurtis L.\ Nusbaum\thanks{St. John's University, klnusbaum@csbsju.edu} \and Dr.\ Mike Heroux\thanks{Sandia National Laboratories,
maherou@sandia.gov}}

\begin{document}

\maketitle

% Include your abstract here.
\begin{abstract}
\input{../OptikaAbstract}
\end{abstract}

\section{Introduction}
In the world of scientific computing there is a problem: most software developers are far more concerned
with the functionality of their software rather than their user interface. This is understandable given
the limited time and pressures of scientific computing environments. And in 
cases where there are only a few users of a piece of software this type of development is tolerable. However, when a piece of
software starts to be used by a wider audience, poor user interface design issues come to the forefront and 
can greatly hinder further adoption of a particular piece of software. Optika\footnote{For more information on Optika, please see
its documentation~\cite{OptikaPackage}} is an attempt to solve this
problem in a generic fashion for parameterized scientific applications.

Since developers of scientific applications don't really care about user interfaces, Optika needs
to provide a minimal amount of hurdles for developers. Also, the end result needs to be an intuitive
user interface that can be easily navigated and utilized regardless of the underlying computations being done.

The purpose of this paper is to discuss the development of the Optika package. In doing so we hope to
demonstrate how Optika solved some of the issues associated with developing a generic user interface
for scientific applications and provide justification for why we chose particular solutions. We will
proceed to discuss Optika development in a semi-chronological fashion.

\input{../OptikaDev}
\appendix
\section{Nomenclature}
\input{../OptikaNomenclature}
\bibliographystyle{siam}
% Edit the line below to be your first and last names.
\bibliography{KurtisNusbaum,../OptikaSANDReport}
% Edit FirstnameLastname below to be your first and last names, but leave the line commented out.
% This line will help me merge bibliographies for the proceedings.
%\input{KurtisNusbaum/KurtisNusbaum.bbl}

\end{document}
