%This turns on the SAND report covers
\newif\ifdraft\draftfalse
\newif\ifsand\sandtrue

\SANDnum{SAND2011-5715}
\SANDauthor{Gregory D. Sjaardema}
\SANDprintDate{\today}
\newcommand{\theTitle}{A Collection of Exodus Utilities:\\Exodiff, EPU, EJoin, and Conjoin}
\title{\theTitle}
\ifsand
\pdfbookmark[1]{Cover}{cover}
\doCover
\newpage
\else
\SANDmarks{cover}
\setcounter{page}{3}
\fi

\pdfbookmark[1]{Title}{title}

%\begin{titlepage}
\begin{center}
\SANDnumVar\\
\SANDreleaseTypeVar\\
\ifdraft
Draft Date: \SANDprintDateVar\\
\else
Printed \SANDprintDateVar\\
\fi

\vspace{0.75in}
\CoverFont{m}{24}{28pt}
\theTitle\\
\vspace{0.75in}
\CoverFont{m}{12}{14pt}
\SANDauthorVar\\
Simulation Modeling Sciences Department\\
Sandia National Laboratories\\
Albuquerque, NM 87185-0380\\
\vspace*{.5in}
\textbf{Abstract}
\end{center}
\vspace{-.4cm}\par

The applications \exodiff{}, \epu{}, \ejoin{}, and \conjoin{} are
members of the Sandia Engineering Analysis Code Access System
(SEACAS~\cite{bib:seacas}) which is used by analysts at Sandia
National Laboratories.  The applications all read and/or write finite
element databases stored in the \exo{}~\cite{bib:exodus} format.

Each application is targeted at a few specific areas of functionality:

\begin{description}
\item[\exodiff] compares the results data from two \exo{} files with
user-specified relative or absolute tolerances.

\item[\epu] combines multiple \exo{} databases produced by a parallel
application into a single \exo{} database.

\item[\ejoin] joins two or more \exo{} databases into a single
\exo{} database. The input databases must have disjoint meta and bulk data.

\item[\conjoin] joins two or more \exo{} databases into a single
database. The input databases should represent the same model geometry
with results data written sequentially.
\end{description}
%\end{titlepage}
