OSKI is a package used to perform optimized sparse matrix-vector operations.
It provides both a statically tuned library created upon installation
and dynamically tuned routines created at runtime.  
OSKI provides support for single and double
precision values of both real and complex types, along with indexing using both integer
and long types.  When possible it follows the sparse BLAS standard \cite{IK:SBLAS}
 as closely as possible in defining operations and functions.  

Before a matrix can use OSKI functionality, it first must be converted to the matrix
type \IKcompfont{oski\_matrix\_t}.  To store a matrix as an \IKcompfont{oski\_matrix\_t} object, a create function
must be called on a CSR or CSC matrix.  An \IKcompfont{oski\_matrix\_t} object can either be created using
a deep or shallow copy of the matrix.  When a shallow copy is created, the user must
only make changes to the matrix's structure through the OSKI interface.  When a deep
copy is created, the matrix that was passed in can be edited by the user as desired.
OSKI automatically makes a deep copy when any matrix is tuned in a manner that changes its structure.

\begin{table}[htbp] 
\begin{center}
\begin{tabular}{|l|l|}
	\hline
Routine & Calculation  \\
	\hline
Matrix-Vector Multiply   & $y = \alpha Ax + \beta y$ or  \\
 & $y = \alpha A^Tx + \beta y$ \\ \hline
Triangular Solve   & $x = \alpha A^{-1}x$ or \\
 & $x = \alpha {A^T}^{-1}x$\\ \hline
Matrix Transpose Matrix-Vector Multiply & $y = \alpha A^TAx + \beta y$ or \\
 & $y = \alpha AA^Tx + \beta y$ \\ \hline
Matrix Power Vector Multiply  & $y = \alpha A^px + \beta y$ or \\
 & $y = \alpha {A^T}^px + \beta y$\\ \hline
Matrix-Vector Multiply and & $y = \alpha Ax + \beta y$ and \\ 
Matrix Transpose Vector Multiply & $z = \omega Aw + \zeta z$ or \\
 & $z = \omega A^Tw + \zeta z$ \\
	\hline
\end{tabular}
\caption{Computational kernels from OSKI available in Epetra.}
\label{IK:fig:oskikernels}
\end{center}
\end{table}

OSKI provides five matrix-vector operations to the user.  The operations are
shown in Table \ref{IK:fig:oskikernels}.  Hermitian operations are available in OSKI,
but are not shown in the table since Epetra does not include Hermitian functionality.  The last three
kernels are composed operations using loop fusion \cite{IK:loopfussion} to increase data reuse.
To further improve performance, OSKI can link to a highly tuned BLAS library.

OSKI creates optimized routines for the target machine's hardware
based on empirical search, in the same manner as ATLAS \cite{IK:ATLAS} and PHiPAC \cite{IK:PHiPAC}.  The goal of the
search is create efficient static kernels to perform the operations listed in Table \ref{IK:fig:oskikernels}.
The static kernels then become the defaults that are called by OSKI when runtime
tuning is not used.  Static tuning can create efficient kernels for a
given data structure.
To use the most efficient kernel, the matrix data structure may need to be reorganized.

When an operation is called enough times to amortize the cost of rearranging the
data structure, runtime tuning can be more profitable than using statically tuned
functions.  OSKI provides multiple ways to invoke runtime tuning, along with multiple
levels of tuning.  A user can explicitly ask for a matrix to always be tuned for a specific
kernel by selecting either the moderate or aggressive tuning option.  If the user wishes for OSKI to 
decide whether enough calls to a function occur to justify tuning, hints can be used.
Possible hints include telling OSKI the number of calls expected to the routine and information
about the matrix, such as block structure or symmetry.  In either
case, OSKI tunes the matrix either according to the
user's requested tuning level, or whether it expects to be able to
amortize the cost of tuning if hints are provided.  Instead of providing hints the user may,
periodically call the tune function.  In this case, the tune function predicts
the number of future kernel calls based on past history, and tunes the routine only
if it expects the tuning cost to be recovered via future routine calls.

OSKI can also save tuning transformations for later reuse.  Thus, the cost of tuning searches can be
amortized over future runs.  Specifically, a search for the best tuning options does not need to be
run again, and only the prescribed transformations need to be applied.  

OSKI is under active development.  As of this writing, the current version is 1.0.1h,
with a multi-core version under development \cite{IK:Vuduc-convo}.
While OSKI provides many optimized sparse matrix kernels,
some features have yet to be implemented, and certain optimizations are missing.
OSKI is lacking multi-vector kernels and stock versions of the composed kernels.
These would greatly add to both OSKI's usability and performance.
The Matrix Power Vector Multiply is not functional.
Finally, OSKI cannot transform (nearly) symmetric matrices to reduce storage or
convert from a CSR to a CSC matrix (or vice versa).  Both could
provide significant memory savings.
Thus, performance gains from runtime tuning should not be expected for point matrices.
An exception is pseudo-random matrices, which may benefit from cache blocking.
