
\cba{export data}{\parameterlist}{Writes a subset of the hierarchy data to files.
      Valid \texttt{string} parameters are 
      ``A'', ``P'', ``R'', ``Nullspace'', and ``Coordinates''.
      The string values are comma separated nonnegative integers within curly braces, e.g.,
      ``\{0,1,3\}''.  Valid \texttt{int} options are \texttt{frequency} and \texttt{total}.
      Setting \texttt{frequency} to a positive integer value \texttt{n} means data is written
      to file on every \texttt{n}th Hierarchy construction.  If \texttt{total} is nonnegative,
      the data will be written at most \texttt{total} times.  If \texttt{total} is negative,
      there is no limit to the number of times the data is written.
      The only valid \texttt{bool} option is \texttt{export first system}.  If true, then writing
      will begin with the first linear system.
      A matrix/multivector with a name ``X'' on level ``L'' is saved in two or three
      three MatrixMarket files: a) data is saved in
      \textit{X\_L.mm}; b) its row map is saved in
      \textit{rowmap\_X\_L.mm}; c) its column map (for matrices) is saved in
      \textit{colmap\_X\_L.mm}.}
          
\cbb{print initial parameters}{bool}{true}{Print parameters provided for a hierarchy construction.}
          
\cbb{print unused parameters}{bool}{true}{Print parameters unused during a hierarchy construction.}
          
\cbb{transpose: use implicit}{bool}{false}{Use implicit transpose for the restriction operator.}
          