
\cbb{aggregation: type}{string}{"uncoupled"}{Aggregation scheme. Possible values: see Table~\ref{t:aggregation}.}
          
\cbb{aggregation: mode}{string}{"uncoupled"}{Controls whether aggregates are allowed to cross processor boundaries. Possible values: "uncoupled" aggregates cannot cross processor boundaries.}
          
\cbb{aggregation: ordering}{string}{"natural"}{Node ordering strategy. Possible values: "natural" (local index order), "graph" (filtered graph breadth-first order), "random" (random local index order).}
          
\cbb{aggregation: phase 1 algorithm}{string}{"Distance2"}{Look at distance2 hops when aggregating.}
          
\cbb{aggregation: symmetrize graph after dropping}{bool}{false}{Symmetrize the graph used for aggregation.}
          
\cbb{aggregation: use blocking}{bool}{false}{Drop connections between different blocks. Blocks are given by a vector "BlockNumber" of scalar type LocalOrdinal that is passed via the "user data" list.}
          
\cbb{aggregation: drop scheme}{string}{"classical"}{Connectivity dropping scheme for a graph used in
      aggregation. Possible values: "point-wise" and "cut-drop". "signed classical sa", "classical", "distance laplacian", "signed classical", "block diagonal", "block diagonal classical", "block diagonal distance laplacian", "block diagonal signed classical", "block diagonal colored signed classical", "signed classical distance laplacian", "signed classical sa distance laplacian" are also allowed, but discouraged. Instead, the corresponding combination of "aggregation: strength-of-connection: matrix", "aggregation: strength-of-connection: measure" and "aggregation: drop scheme" should be used.}
          
\cbb{aggregation: strength-of-connection: matrix}{string}{"A"}{The matrix that is used for constructing the graph used for aggregation. Possible values: "A" (meaning the level matrix itself) and "distance laplacian".}
          
\cbb{aggregation: strength-of-connection: measure}{string}{"smoothed aggregation"}{The strength measure used for constructing the graph. Possible values: "smoothed aggregation", "signed smoothed aggregation", "signed ruge-stueben", "unscaled".}
          
\cbb{aggregation: distance laplacian metric}{string}{unweighted}{Metric used to compute the distance Laplacian. Possible values: "unweighted", "material"}
          
\cbb{aggregation: drop tol}{double}{0.0}{Connectivity dropping threshold for a graph used in aggregation.}
          
\cbb{aggregation: use ml scaling of drop tol}{bool}{false}{Enables ML-style scaling of drop tol, where the drop tol halves with each successive level.}
          
\cbb{aggregation: min agg size}{int}{2}{Minimum size of an aggregate.}
          
\cbb{aggregation: max agg size}{int}{-1}{Maximum size of an aggregate (-1 means unlimited).}
          
\cbb{aggregation: compute aggregate qualities}{bool}{false}{Whether to compute aggregate quality estimates.}
          
\cbb{aggregation: brick x size}{int}{2}{Number of points for x axis in "brick" aggregation (limited to 3).}
          
\cbb{aggregation: brick y size}{int}{2}{Number of points for y axis in "brick" aggregation (limited to 3).}
          
\cbb{aggregation: brick z size}{int}{2}{Number of points for z axis in "brick" aggregation (limited to 3).}
          
\cbb{aggregation: brick x Dirichlet}{bool}{false}{Asserts that Dirichlet conditions are applied in
        the x-direction and the Dirichlet DOFs are not aggregated.}
          
\cbb{aggregation: brick y Dirichlet}{bool}{false}{Asserts that Dirichlet conditions are applied in
        the y-direction and the Dirichlet DOFs are not aggregated.}
          
\cbb{aggregation: brick z Dirichlet}{bool}{false}{Asserts that Dirichlet conditions are applied in
        the z-direction and the Dirichlet DOFs are not aggregated.}
          
\cbb{aggregation: Dirichlet threshold}{double}{0.0}{Threshold for determining whether entries are zero during Dirichlet row detection.}
          
\cbb{aggregation: greedy Dirichlet}{bool}{false}{Force the aggregate to be Dirichlet if any DOFs in the aggregate are Dirichlet (default is aggregates are Dirichlet only if all DOFs in the aggregate are Dirichlet).}
          
\cbb{aggregation: deterministic}{bool}{false}{Boolean indicating whether or not aggregation will be run deterministically in the kokkos refactored path (only used in uncoupled aggregation).}
          
\cbb{aggregation: coloring algorithm}{string}{serial}{Choice of distance 2 independent set or coloring algorithm used by Uncoupled Aggregation, when using kokkos refactored aggregation. See Table \ref{t:coloring_algs} for more information.}
          
\cbb{aggregation: dropping may create Dirichlet}{bool}{true}{If true, any matrix row has nonzero off-diagonal entries will be treated as Dirichlet if aggregation dropping leaves only the diagonal entry.}
          
\cbb{aggregation: export visualization data}{bool}{false}{Export data for visualization post-processing.}
          
\cbb{aggregation: output filename}{string}{""}{Filename to write VTK visualization data to.}
          
\cbb{aggregation: output file: time step}{int}{0}{Time step ID for non-linear problems.}
          
\cbb{aggregation: output file: iter}{int}{0}{Iteration for non-linear problems.}
          
\cbb{aggregation: output file: agg style}{string}{Point Cloud}{Style of aggregate visualization.}
          
\cbb{aggregation: output file: fine graph edges}{bool}{false}{Whether to draw all fine node connections along with the aggregates.}
          
\cbb{aggregation: output file: coarse graph edges}{bool}{false}{Whether to draw all coarse node connections along with the aggregates.}
          
\cbb{aggregation: output file: build colormap}{bool}{false}{Whether to output a random colormap in a separate XML file.}
          
\cbb{aggregation: output file: aggregate qualities}{bool}{false}{Whether to plot the aggregate quality.}
          
\cbb{aggregation: output file: material}{bool}{false}{Whether to plot the material.}
          
\cbb{aggregation: mesh layout}{string}{Global Lexicographic}{Type of ordering for structured mesh aggregation. Possible values: "Global Lexicographic" and "Local Lexicographic".}
          
\cbb{aggregation: output type}{string}{Aggregates}{Type of object holding the aggregation data. Possible values: "Aggregates" or "CrsGraph".}
          
\cbb{aggregation: coarsening rate}{string}{{3}}{Coarsening rate per spatial dimensions, the string must be interpretable as an array by Teuchos.}
          
\cbb{aggregation: number of spatial dimensions}{int}{3}{The number of spatial dimensions in the problem.}
          
\cbb{aggregation: coarsening order}{int}{0}{The interpolation order used while constructing these aggregates, this value will be passed to the prolongator factory. There, possible values are 0 for piece-wise constant and 1 for piece-wise linear interpolation to transfer values from coarse points to fine points. }
          