
\cbb{repartition: enable}{bool}{false}{Rebalancing on/off switch.}
          
\cbb{repartition: partitioner}{string}{"zoltan2"}{Partitioning package to use. Possible values: "zoltan" (\zoltan{} library), "zoltan2" (\zoltantwo{} library).}
          
\cba{repartition: params}{\parameterlist}{Partitioner parameters. \muelu passes them directly to the appropriate package library. In particular, this allows to choose a partitioning algorithm from \zoltan{} or \zoltan2{} or from external packages such as \parmetis{}.}
          
\cbb{repartition: start level}{int}{2}{Minimum level to run partitioner. \muelu does not rebalance levels finer than this one.}
          
\cbb{repartition: min rows per proc}{int}{800}{Minimum number of rows per MPI process. If the actual number if smaller, then rebalancing occurs. The value is not used if "repartition: min rows per thread" is positive.}
          
\cbb{repartition: target rows per proc}{int}{0}{Target number of rows per MPI process after rebalancing. If the value is set to 0, it will use the value of "repartition: min rows per proc"}
          
\cbb{repartition: min rows per thread}{int}{0}{Minimum number of rows per thread. If the actual number if smaller, then rebalancing occurs. If the value is set to 0, no repartitioning based on thread count will occur.}
          
\cbb{repartition: target rows per thread}{int}{0}{Target number of rows per thread after rebalancing. If the value is set to 0, it will use the value of "repartition: min rows per thread".}
          
\cbb{repartition: max imbalance}{double}{1.2}{Maximum nonzero imbalance ratio. If the actual number is larger, the rebalancing occurs.}
          
\cbb{repartition: remap parts}{bool}{true}{Postprocessing for partitioning to reduce data migration.}
          
\cbb{repartition: rebalance P and R}{bool}{false}{Explicit rebalancing of R and P during the setup. This speeds up the solve, but slows down the setup phases.}
          
\cbb{repartition: explicit via new copy rebalance P and R}{bool}{false}{Fully explicit rebalancing of R and P during setup that makes new copy and invokes fill copy.  Slowest rebalancing option, but needed by combine mg algo.}
          