\section*{Nomenclature}
\addcontentsline{toc}{section}{Nomenclature}
\begin{itemize}

\item[Package]
A collection of software focused on one primary class of numerical
methods.  Also a fundamental, integral unit in the Trilinos framework.

\item[Trilinos]
The name of the project.  Also a Greek term which,
loosely translated means ``a string of pearls,'' 
meant to evoke an image that each Trilinos package is a pearl in its 
own right, but is even more valuable when combined with other 
packages.

\item[Didasko]
The tutorial of Trilinos.  Offers a quick introduction to several
Trilinos packages. It contains a printable PDF document and a
variety of well-commented browsable examples that illustrate how
to use Trilinos.

Didasko contains examples for the following Trilinos packages: 
Teuchos, Epetra, EpetraExt, TriUtils, Galeri, AztecOO, IFPACK,ML,
NOX, Anasazi, TPetra.

\item[new\_package] A sample Trilinos package containing all of the
infrastructure to install a new package into the Trilinos framework.
Contains the basic directory structure, a collection of sample
configuration and build files and a sample ``Hello World'' package.

\item[Amesos]
The Direct Sparse Solver Package in Trilinos. The goal of
Amesos is to make AX=B as easy as it sounds, at least for direct
methods. Amesos provides clean and consistent interfaces to 
several popular third party libraries.

\item[Anasazi]
An extensible and interoperable framework for large-scale eigenvalue
algorithms.The motivation for this framework is to provide a generic
interface to a collection of algorithms for solving large-scale 
eigenvalue problems.

\item[AztecOO] 
Linear solver package based on preconditioned Krylov methods.  A 
follow-on to the Aztec solver package~\cite{Aztec2.1}.  
Supports all Aztec 
interfaces and functionality, but also provides significant new 
functionality.

\item[Belos] A Greek term meaning ``arrow.'' Belos is the next
generation of iterative solvers.  Belos solvers are written using
``generic'' programming techniques.  In other words, Belos is written
using TSF abstract interfaces and therefore has no explicit dependence
on any concrete linear algebra library.  Instead, Belos solvers can be
used with any concrete linear algebra library that implements the TSF
abstract interfaces. 

\item[Claps]
Claps is a collection of domain decomposition preconditioners and
solvers.

\item[Ifpack] 
Object-oriented algebraic preconditioner, compatible with 
Epetra and AztecOO.  Supports construction and use of parallel
distributed memory preconditioners such as overlapping Schwarz domain
decomposition, Jacobi scaling and local Gauss-Seidel relaxations.

\item[Komplex] 
Complex linear equation solver using equivalent real 
formulations~\cite{DayHero2000}, built on top of Epetra and AztecOO.

\item[Meros]
Segregated preconditioning package.  Provides scalable block
preconditioning for problems that couple simultaneous solution
variables such as Navier-Stokes problems.

\item[ML]
Algebraic multi-level preconditioner package.  Provides scalable
preconditioning capabilities for a variety of problem classes.

\item[NOX]
A collection of nonlinear solvers, designed to be easily integrated
into an application and used with many different linear solvers.

\item[Petra]
A Greek term meaning ``foundation.''  Trilinos has three Petra 
libraries: Epetra, Tpetra and Jpetra that provide basic classes 
for constructing and manipulating matrix, graph and vector
objects.  Epetra is the current production version that is
split into two packages, one core and one extensions.

\item[TSF]
Composed of several packages.  Thyra provides 
a basic collection of abstract interfaces to vectors, linear
operators, solvers, etc.  These interfaces provide a common
interface for applications to access one or more packages that
implement the abstract interface.  These interfaces can also be used
by other packages in Trilinos to accomplish the same purpose.
TSF Extended builds on top of Thyra, providing implicit
aggregation capabilities and overloaded operators.

\end{itemize}

