\chapter{Command Summary} \label{a:cmdsum}

\section{Operational Commands
(page~\protect\pageref{sec:oper})}\label{sec:opsum} 

\cmdsum{\cmd{MASS} 
      \param{nquad},
      \optparam{density}\\
\cmd{PROPERTIES} 
      \param{nquad} 
      \optparam{density}
} {
calculates several mass properties of the body. 
} 

\cmdsum{\cmd{LOCATE} \cmd{\{NODES$|$ELEMENTS\}} 
      \param{locate\_option}
} {
locates \cmd{NODES} or \cmd{ELEMENTS} that are a specified distance from
a user-defined point, line, or plane. 

\begingroup\small
\tabcolsep=3pt
\begin{tabular}{*{3}{l}  *{6}{l}  *{2}{l} }
LOCATE & NODES$|$ELEMENTS & POINT & $x_0$ & $y_0$ & $z_0$ 
                                  &       &       &       &distance &toler \\
LOCATE & NODES$|$ELEMENTS & LINE  & $x_0$ & $y_0$ & $z_0$ &
                                    $x_1$ & $y_1$ & $z_1$ &distance &toler \\
LOCATE & NODES$|$ELEMENTS & PLANE & $x_0$ & $y_0$ & $z_0$ &
                                     $i$  &  $j$  &  $k$  &distance &toler \\
\end{tabular}
\endgroup
}

\cmdsum{\cmd{CAVITY  } 
   \param{sset$_{1}$}, \param{sset$_{2}$}, \ldots, \param{sset$_n$},
   \optparam{\cmd{CENTER}}, \optparam{xcen}, \optparam{ycen},
   \optparam{zcen}
} {
calculates the volume of a cavity defined by the element
sidesets \param{sset$_{1}$} to \param{sset$_n$}.  
}

\cmdsum{\cmd{LIMITS} \optparam{\cmd{ALLTIMES}}
} {
determines the minimum, maximum, and range of the 
X, Y, and Z coordinates for each of the selected material blocks.
}

\cmdsum{\cmd{OVERLAP }
      \param{flag$_m$}, \param{flag$_s$}
} {
determines whether any of the nodes on the surface defined by
the sideset \param{flag$_s$} penetrate the element faces on the surface
defined by the sideset \param{flag$_m$}.  
}

\cmdsum{\cmd{TIMESTEP}
      \param{wave\_speed} 
      \optparam{damping\_fraction} 
} {
estimates the maximum stable time step for an explicit transient
dynamics finite element program. 
}

\cmdsum{\cmd{GAP} 
      \param{flag$_m$}, 
      \param{flag$_s$}, 
      \param{d$_{\max}$}, 
      \optparam{\cmd{DISTANCE$|$NORMAL}}
} {
calculates the distance between nodes on the surfaces defined
by the sideset flags \param{flag$_m$} and \param{flag$_s$}.  
}

\newpage
\section{Parameter Setting Commands (page~\protect\pageref{sec:param})}

\cmdsum{\cmd{AXISYMMETRIC} 
} {
the two-dimensional body is axisymmetric.  
}

\cmdsum{\cmd{PLANAR} or \cmd{PLANE}
} {
the two-dimensional body described is planar or plane strain. 
}

\cmdsum{\cmd{EXODUS} 
      \optparam{\cmd{ON$|$OFF}} 
} {
indicates that the operational commands should calculate values based on
the undeformed geometry (\cmd{off}), or for every time step (\cmd{on}). 
}

\cmdsum{\cmd{DENSITY }
} {
prompts the user for the density of each of the element blocks 
}

\cmdsum{\cmd{SORT}
   \param{sort\_field}
   \optparam{sort\_order}
} {
orders the output of a \cmd{LOCATE} command according to
\param{sort\_field} and \param{sort\_order}. 
}

\cmdsum{\cmd{SELECT}
   \param{option} 
} {
selects the materials or material blocks for the operational commands. 
}

\cmdsum{\cmd{SELECT BLOCKS} 
   \optparam{\cmd{ALL},}
   \param{block\_id$_{1}$}, \param{block\_id$_{2}$}, \ldots\
} {
selects the element blocks. 
} 

\cmdsum{\cmd{SELECT MATERIALS} 
   \optparam{\cmd{ALL},}
   \param{material\_id$_{1}$}, \param{material\_id$_{2}$}, \ldots\
} {
selects the element blocks with the material number corresponding to
\param{material\_id}.  
}

\cmdsum{\cmd{TMIN}
   \param{tmin}
} {
sets the minimum selected time to \param{tmin}.
}

\cmdsum{\cmd{TMAX}
   \param{tmax}
} {
sets the maximum selected time to \param{tmax}.
}

\cmdsum{\cmd{DELTIME}
   \param{delt}
} {
sets the selected time interval to \param{delt}.
}

\cmdsum{\cmd{TIMES}
   \param{tmin} \cmd{TO} \param{tmax} \cmd{BY} \param{delt}
} {
sets \param{tmin}, \param{tmax}, and \param{delt} with a single command. 
}

\section{Database Display Commands (page~\protect\pageref{cmd:display})}

\cmdsum{\cmd{LIST}
   \param{option} 
} {
displays the database information specified by \param{option}.

\cenlinesbegin
\cmd{LIST \{VARS$|$VARIABLES\}}\\
\cmd{LIST \{BLOCKS$|$MATERIALS\}}\\
\cmd{LIST \{NSETS$|$NODESETS\}}\\
\cmd{LIST \{SSETS$|$SIDESETS\}}\\
\cmd{LIST STEPS}\\
\cmd{LIST TIMES}\\
\cenlinesend
}

\cmdsum{\cmd{HELP}
   \param{command}
} {
displays information about a program command.
}

\section{Miscellaneous Commands (page~\protect\pageref{sec:misc})}

\cmdsum{\cmd{ECHO}
      \optparam{\cmd{ON$|$OFF}}
} {
controls the output of results to the terminal.  
}

\cmdsum{\cmd{PRINT}
      \optparam{\cmd{ON$|$OFF}} 
} {
controls the output of results to the output file.  
}

\cmdsum{\cmd{COMMENT }   
      \optparam{\cmd{PAGE},}
      \param{ncom}
} {
prompts the user for \param{ncom} comment lines which are written to
the output file. 
}

\cmdsum{\cmd{EXIT    } 
} {
exits the program and closes all files.
}

