\documentclass{article}
\title{TivaBuena Information}
\author{Kurtis Nusbaum}
\date{May 18th 2009}
\begin{document}
\maketitle

\section{Introduction}
The TivaBuena package is intended to give Trilinos Developers an easy way to construct a GUI for their program. Through the use of Teuchos ParameterLists,
the TivaBuena package will allow the developer to achieve the following workflow within their program:
	\begin{enumerate}
	\item Specify a list of inputs needed from the user
	\item Define valid values for inputs as well as defaults
	\item Present the required inputs to the user in the form of a GUI
	\item Allow the user to change the values of the inputs to their liking
	\item Obtain the values the user entered in the form of a Parameter List.
	\end{enumerate}

\subsection{Compile Notes}
In order to compile TivaBuena you must have Qt installed. You can either download Qt directly from the website (http://www.qtsoftware.com) or if you are on Linux, you can use your package manager.
If you use your package manager, may sure you download the Qt developement package (usually something like qt4-devel) and that it is Qt version 4.3 or higher. If you download Qt directly and put it some where other than the default
	installation directory you will have to compile trilinos using these cmake arguments:
	\begin{itemize}
	\item -DQT\_INCLUDE\_DIRS:STRING=directory containing your qt include files
	\item -DQT\_LIBRARY\_DIRS:STRING=directory containing your qt libraries
	\item -DQT\_BINARY\_DIR:STRING=directory containing your qt binaries
	\end{itemize}	

\section{Contact}

If you are having any problems with StratRunner, please contact Kurtis Nusbaum at klnusbaum@csbsju.edu.

\section{Acknoledgements}

I would like to thank Dr. Mike Heroux for giving me the wonderful opportunity to do reasearch for him. It has been a truely enjoyable experience. I 
would like to thank Jonathan Hu and Christopher Baker for there help in determining options for certain parameters. Lastly I would like to thank all
my Family, Friends, and specifically my Mother and Father. You have all been so amazing and there are not words to express my gratitued towards you all.
My love for you is infinite.
\end{document}
