
\section{ParameterList Description}

The following description of the Rythmos ParameterList was auto-generated
from the ParameterList itself, thus allowing easy refreshing of the
input description and providing up-to-date information. Fig.~\ref{fig:ParameterList-schematic}
is a hyperlinked, indented heirarchy of the Rythmos ParameterList,
that provides an overall structure of the Rythmos ParameterList. The
hyerlinks are connected to subsequent sections, which contain additional
hyperlinks to parent and children sublists, and descriptions of the
sublist and parameters. All these hyperlinks provide an easy way to
hop around the parameters in order to setup the time-integrator. 

The basic structure starts with the ``Integrator Base'', and is
broken down into the Integrator and the Stepper. Each of these contain
several sublists giving the details of the Integrator and Stepper,
e.g., IntegrationControlStrategy, InterpolationBuffer, StepControlStrategy,
and error weight vector.

An easy way to create a Rythmos ParameterList is to get a valid ParameterList
from the IntegatorBuilder, and manipulate the parameters from there.
In this way, one able to correctly setup the Integrator and Stepper
for a transient soution. For example, \input{RythmosParameterListExample.tex}
Obviously, most software will get its input specification from the
XML version of the ParameterLists, and then pass them onto Rythmos.
The description provided here can also help in setting up those XML
files. 

\input{RythmosUsersManualParameterList_Raw.tex}
